### 2020-05-14 paper2sattose.tex
creation

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolceurws}

\title{Damegender: Writing and Comparing Gender Detection Tools}

\author{
David Arroyo Menéndez \\ GSYC \\ Universidad Rey Juan Carlos I \\ d.arroyome@alumnos.urjc.es
\and
Jesús González Barahona \\ GSYC \\ Universidad Rey Juan Carlos I \\ jgb@gsyc.es
}

\institution{}

\begin{document}
\maketitle

\begin{abstract}
The variable sex (male or female) is one of most used variables for
any study in sociology, but this variable can be hidden in Internet
communities. The gender detection from a name is an important problem
in Natural Language Processing to decide if a string labeled as name
is classified as male or female. An engineer will find useful
make gender detection from a name retrieving information from social
networks, mailing lists, instant messaging, software repositories,
papers, etc. To achieve gender equality and empower all women and
girls is a goal in sustanaible development in United Nations, so to
measure the gender gap is a previous step to find solutions to reduce
it.

Nowadays, there are several Application Programming Interfaces to
guess gender from a name. This kind of software has the database
based on propietary databases and the software is not free, so some
scientific works are difficult to reproduce.

In this paper, we are envisioning how to solve these problems,
offering a solution with a free license and open data names from
official census useful to replace, use and/or compare these apis with
very good results. This tool provides Machine Learning to predict
strings, that's useful to guess diminutives or nicknames.
\end{abstract}

\section{Introduction}

There are different ways to detect gender from a person name and
perhaps a surname: census, wikipedia, self-references in trust
websites, ... The most common way to detect gender from a name is the
Application Programming Interfaces with a good popularity, for
example, genderapi, namsor, genderize, ...

The problems addressed are:
\begin{itemize}
\item Evaluate quality/price with different commercial solutions.
\item Think about solutions using free licenses.
\item Treatment with names without census, for example, nicknames,
  diminutives, ...
\item Massive gender detection from Internet, for example, mailing
  lists, software repositories, ...
\end{itemize}

In this paper, these problems are faced writing a Python solution for:

\begin{itemize}
\item To evaluate quality of different solutions applying metrics
  suggested by ~\cite{10.7717/peerj-cs.156}
\item To understand the current technology in detail, I have developed
  a tool guessing gender from a name giving support to Spanish and
  English from the open data census provides by the states.
\item To fix the problem with nicknames and diminutives, we have
  developed a machine learning solution to strings not found in the
  census dataset.
\item To do proof-of-concept tests applying Perceval to detect
  gender in mailing lists and software repositories.
\end{itemize}

\section{First Level Heading}

First level headings are all flush left, initial caps, bold and in point
size 12. One line space before the first level heading and $1/2$ line
space after the first level heading.

\subsection{Second Level Heading}

Second level headings must be flush left, initial caps, bold and in point
size 10. One line space before the second level heading and $1/2$ line
space after the second level heading.

\subsubsection{Third Level Heading}

Third level headings must be flush left, initial caps and bold.
One line space before the third level heading and $1/2$ line
space after the third level heading.

\paragraph{Fourth Level Heading}

Fourth level headings must be flush left, initial caps and roman type.
One line space before the fourth level heading and $1/2$ line
space after the fourth level heading.

\subsection{Citations In Text}

Citations within the text should indicate the author's last name and
year\cite{Knuth-vol3}. Reference style\cite{Comer-btree}
should follow the style that you are used to using, as long as the
citation style is consistent.

\subsubsection{Footnotes}

Indicate footnotes with a number\footnote{This is a sample footnote} in
the text. Place the footnotes at the bottom of the page they appear on.
Precede the footnote with a vertical rule of 2 inches (12 picas).

\subsubsection{Figures}

All artwork must be centered, neat, clean and legible. Do not use pencil
or hand-drawn artwork. Figure number and caption always appear after the
the figure. Place one line space before the figure, one line space
before the figure caption and one line space after the figure caption.
The figure caption is initial caps and each figure is numbered
consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page to avoid
splitting the figure and figure caption.

Figure \ref{fig1} shows how to include a figure as encapsulated postscript.
The source of the figure is in file {\tt fig1.eps}.

\begin{figure}[ht]
\begin{center}
\includegraphics[height=4cm]{fig1}
\caption{Sample EPS figure }
\label{fig1}
\end{center}
\end{figure}

Below is another figure using LaTeX commands.

\begin{figure}[ht]
\begin{center}
\setlength{\unitlength}{1pt}
\footnotesize
\begin{picture}(160,80)
        \put(0,0){\framebox(160,80)[]{}}
        \put(10,35){\framebox(80,40){}}
        \put(100,20){\framebox(40,20){}}
        \put(70,10){\framebox(20,10){}}
        \put(20,5){\framebox(10,5){}}
\end{picture}
\caption{Sample Figure Caption}
\end{center}
\end{figure}

\subsubsection{Tables}

All tables must be centered, neat, clean and legible. Do not use pencil
or hand-drawn tables. Table number and title always appear before the
table.

One line space before the table title, one line space after the table
title and one line space after the table. The table title must be
initial caps and each table numbered consecutively.

\begin{table}[ht]
\begin{center}
\caption{Sample Table}

\bigskip

\begin{tabular}{|l|l|r|}
\hline
A & B & 1\\ \hline
C & D & 2\\
E & F & 3\\ \hline
\end{tabular}
\end{center}
\end{table}

\subsubsection{Handling References}

Use a first level heading for the references. References follow the
acknowledgements.

\subsubsection{Acknowledgements}

Use a third level heading for the acknowledgements. All acknowledgements
go at the end of the paper.

%\bibliographystyle{alpha} 
%\bibliography{samplebib}
%inline the .bbl file directly for mailing to authors.

\begin{thebibliography}

\bibitem[Com79]{Comer-btree}
D.~Comer.
\newblock The ubiquitous b-tree.
\newblock {\em Computing Surveys}, 11(2):121--137, June 1979.

\bibitem[Knu73]{Knuth-vol3}
Santamaría, Lucía and Mihaljević, Helena
\newblock {\em Comparison and benchmark of name-to-gender inference services}.
\newblock Addison-Wesley, 1973.

\end{thebibliography}

\end{document}

### 2020-05-14 sattose.tex
creation

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{twocolceurws}

\title{Damegender: Writing and Comparing Gender Detection Tools}

\author{
David Arroyo Menéndez \\ GSYC \\ Universidad Rey Juan Carlos I \\ d.arroyome@alumnos.urjc.es
\and
Jesús González Barahona \\ GSYC \\ Universidad Rey Juan Carlos I \\ jgb@gsyc.es
}

\institution{}

\begin{document}
\maketitle

\begin{abstract}
The variable sex (male or female) is one of most used variables for
any study in sociology, but this variable can be hidden in Internet
communities. The gender detection from a name is an important problem
in Natural Language Processing to decide if a string labeled as name
is classified as male or female. An engineer will find useful
make gender detection from a name retrieving information from social
networks, mailing lists, instant messaging, software repositories,
papers, etc. To achieve gender equality and empower all women and
girls is a goal in sustanaible development in United Nations, so to
measure the gender gap is a previous step to find solutions to reduce
it.

Nowadays, there are several Application Programming Interfaces to
guess gender from a name. This kind of software has the database
based on propietary databases and the software is not free, so some
scientific works are difficult to reproduce.

In this paper, we are envisioning how to solve these problems,
offering a solution with a free license and open data names from
official census useful to replace, use and/or compare these apis with
very good results. This tool provides Machine Learning to predict
strings, that's useful to guess diminutives or nicknames.
\end{abstract}

\section{Introduction}

There are different ways to detect gender from a person name and
perhaps a surname: census, wikipedia, self-references in trust
websites, ... The most common way to detect gender from a name is the
Application Programming Interfaces with a good popularity, for
example, genderapi, namsor, genderize, ...
~\cite{10.7717/peerj-cs.156}

The problems addressed are:
\begin{itemize}
\item Evaluate quality/price with different commercial solutions.
\item Think about solutions using free licenses.
\item Treatment with names without census, for example, nicknames,
  diminutives, ...
\item Massive gender detection from Internet, for example, mailing
  lists, software repositories, ...
\end{itemize}

In this paper, these problems are faced writing a Python solution for:
\begin{itemize}
\item To evaluate quality of different solutions applying metrics
  suggested by ~\cite{10.7717/peerj-cs.156}
\item To understand the current technology in detail, I have developed
  a tool guessing gender from a name giving support to Spanish and
  English from the open data census provides by the states.
\item To fix the problem with nicknames and diminutives, we have
  developed a machine learning solution to strings not found in the
  census dataset.
\item To do proof-of-concept tests applying Perceval to detect
  gender in mailing lists and software repositories.
\end{itemize}

In Section 2, we explain the current solutions to the problems. In
Section 3, we present the results evaluating the current Application
Programming Interfaces with our software. In Section 4, we discuss
attempts and problems releasing with a free license a gender detection
from name program. In Section 5, we discuss how to obtain Open Datasets
counting names and gender. In Section 6, we describe our
machine learning solution. In Section 7, we describe general
implementation details. Finally, in Section 8 we summarize our
findings, and describe extensions to the work that we are currently
exploring.

\section*{3. Application Programming Interfaces}

\subsection*{Market}

We have reproduced to ~\cite{10.7717/peerj-cs.156} and updated on
27/06/2019 and we are showing the results in
~\ref{table:featuresAndGenderDetectionToolsByName}.

\begin{longtable}[]{@{}lllllll@{}}
\toprule
Feature & Gender API & genderguesser & genderize.io & NameAPI & NamSor & Damegender\tabularnewline
\midrule
\endhead
Database size & 431*10^6 & 48.528 & 114*10^6 & 1.428.345 & 4407*10^6  & 197.271\footnote{This cipher is the sum of names in United States of America, United Kingdom, Uruguay and Spain} \tabularnewline
Regular data updates & yes & no & yes & yes & yes & yes, dev\tabularnewline
Unstructured strings & yes & no & no & yes & no & yes\tabularnewline
Handles surnames & yes & no & no & yes & yes & yes\tabularnewline
Non-Latin alphabets & partially & no & partially & yes & yes & no\tabularnewline
Geo-localization & yes & no & no & yes & yes & no\tabularnewline
Exists locale & yes & yes & yes & yes & yes & yes\tabularnewline
Assingment type & probabilistic & binary & probabilistic & probabilistic & probabilistic & prob\tabularnewline
Free params & total, prob & gender & total, prob & confidence & scale & total, prob\tabularnewline
Guessing with ML & no & no & no & no & no & yes\tabularnewline
Free license & no & yes & no & no & no & yes\tabularnewline
API & yes & no & yes & yes & yes & future\tabularnewline
free requests limited & yes (200) & unlimited & yes (1000) & yes & yes & unlimited\tabularnewline
\bottomrule
\caption{Features and gender detection tools by name}
\label{table:featuresAndGenderDetectionToolsByName}
\end{longtable}

All solutions have increased the database size from
~\cite{10.7717/peerj-cs.156}. NameAPI and GenderAPI is reaching more
features. The tools with a free license have not many features, so for
now that will not be the trend in many situations. Today, one good
solution quality and price is Namsor, which provides unlimited names
through an Application Programming Interface with many features in the
task to detect gender from the name.

\subsection*{Reproducing accuracies and confusion matrix}

~\cite{10.7717/peerj-cs.156} explains different ways to determine
gender from a name by humans and it gives 7000 names applying these
methods. In this dataset the gender is classified as male, female or
unknown. We have used this dataset, but only male and female to these
experiments. We are showing the results in the next table:

\begin{longtable}[]{@{}lllllll@{}}
  \toprule
  API & Accuracy & Precision & F1score & Recall\tabularnewline
\midrule
\endhead
Genderapi & 0.9687686966482124 & 0.9717050018254838 & 0.9637877964874163 & 1.0\tabularnewline
Genderize & 0.926775 & 0.9761303240374678 & 0.9655113956503119 & 1.0\tabularnewline
Damegender (SVC)\footnote{SVC is the acronym of Support Vector Classification, the Machine Learning algorithm that Damegender was using with this results} & 0.8791969539633091 & 0.9718767935718385 & 0.9718767935718385 & 1.0\tabularnewline
Namsor & 0.8672551055728626 & 0.9730097087378641 & 0.9236866359447006 & 1.0\tabularnewline
Nameapi & 0.8301886792452831 & 0.97420272191753 & 0.9054181612233341 & 1.0\tabularnewline
Gender Guesser & 0.7743554248139817 & 0.9848151408450704 & 0.8715900233826968 & 1.0\tabularnewline
\bottomrule
\caption{Different accuracies measures}
\label{table:DifferentAccuraciesMeasures}
\end{longtable}

In ~\ref{table:DifferentAccuraciesMeasures} Genderapi and Genderize
are obtaining the best results, although all solutions is reaching
results better than 0.8 except Gender Guesser.

\begin{longtable}[]{@{}lllll@{}}
  \toprule
  APIs          & gender & male & female & undefined \tabularnewline
\midrule
\endhead
Genderapi         & male    & 3589 & 155  &  67 \tabularnewline
                  & female  & 211  & 1734 &  23 \tabularnewline
Damegender (SVC)\footnotemark[1]  & male    & 3663 & 147  &   0 \tabularnewline
                  & female  & 551  & 1497 &   0 \tabularnewline
Genderguesser     & male    & 3326 &  139 & 346 \tabularnewline
                  & female  & 78   & 1686 & 204 \tabularnewline
Namsor            & male    & 3325 & 139  & 346 \tabularnewline
                  & female  & 78   & 1686 & 204 \tabularnewline
Genderize         & male    & 3157 & 242  & 412 \tabularnewline
                  & female  & 75   & 1742 & 151 \tabularnewline
Nameapi           & male    & 2627 & 674  & 507 \tabularnewline
                  & female  & 667  & 1061 & 240 \tabularnewline 
\bottomrule
\caption{Confusion matrix tables by APIs}
\label{table:ConfusionMatrixTables}
\end{longtable}

With Damegender has been done a comparison about confusion matrix
tables depending the API (see ~\ref{table:ConfusionMatrixTables}). If
we compare these results with the results obtained in
~\cite{10.7717/peerj-cs.156}, we can understand that the results are
similar.

Genderapi has similar results, but it is being improved the undefined
results. In Genderguesser is we are obtaining different results and it
is strange, because the software has not modified from some years
ago. In Genderize we are obtaining the same results. In Nameapi the
guessed results is changing from male to female with more errors. In
Namsor the results is so similar. Damegender is not guessing undefined
because we predict with machine learning if the string is not in the
database.

The most important tools Namsor, Genderapi and Genderize are improving the
accuracies with respect the previous comparison.

\begin{longtable}[]{@{}lllll@{}}
\toprule
API & error code & error code without na & na coded & error gender bias\tabularnewline
\midrule
\endhead
Damegender (SVC)\footnotemark[1] & 0.121 & 0.121 & 0.0 & -0.07\tabularnewline
GenderApi & 0.167 & 0.167 & 0.0 & -0.167\tabularnewline
Gender Guesser & 0.225 & 0.027 & 0.204 & 0.003\tabularnewline
Genderize & 0.276 & 0.261 & 0.0204 & -0.0084 \tabularnewline 
Namsor & 0.332 & 0.262 & 0.095 & 0.01 \tabularnewline
Nameapi & 0.361 & 0.267 & 0.129 & 0.001 \tabularnewline
\bottomrule
\caption{APIs and Errors}
\label{table:ApisAndErrors}
\end{longtable}

In the table it is possible to observe a high index of errors in
Nameapi and Namsor and a low index of errors in GenderApi
and Damegender.

\section*{Datasets}

We can divide the next options choosing a dataset: (1) a census
published with a free license (open census way), (2) a dataset done by
scientist with a paper in a magazine (scientific way), (3) a dataset
released with a free license in a free software package (free software
way), (4) a dataset retrieved from commercial Application Programming
Interfaces (commercial api way).

\begin{verbatim}
$ python3 main.py David --total="ine"
David gender is male
363559  males for David from INE.es
0 females for David from INE.es
\end{verbatim}

In Damegender, we are including Open Data census about names and
gender, such as INE.es or USA and United Kingdom (births and dies). We
want datasets provided by the software package to increment the speed
retrieving data.

From the user final point of view, the value of using Open Data is
give a good explanation when we are asking about the gender from a
name (number of males and females using a specific name in a country)
versus a probability created by the way explained in
~\cite{10.7717/peerj-cs.156} or similar.

From the scientific point of view, the value of using Open Data is to
allow that the experiment can be reviewed by peers on an automatic and
legal way (using proprietary data the reviewer should request it
separately to make the review).

A second approach is to build the dataset reviewing the names in
scientific personal sites, Wikipedia,
... ~\cite{10.7717/peerj-cs.156}. This approach is valid, but it
consumes many time and efforts, although could be useful if there not
a legal way to build the dataset.

A third approach is using a dataset from a popular free software
solution. For instance, Natural Language Tool Kit is providing 8000
labeled english names. The classification is male or female. The
problem again is about don't retrieve data with the social science
quality of National Statistics Institutes. Another example is Gender
Guesser a good dataset for international names with different
categories to define the probability. This approach is similar to use
a dataset released with a paper in a journal, the advantage is to
understand and add new names with a solid criteria accepted by the
scientific community.

We are using the census approach as base of truth to distinguish if a
name is male or female in a geographical area. Generally, a name has a
strong weight to determine if it's a male or a female on this way, for
instance, David is registered 365196 times as male and 0 times as
female in Spain National Institute of Statistics.

Many countries don't provide Open Data census about gender and names,
but we envisioned build a Dataset about names and gender free and
universal working from Gender Guesser dataset and Wikidata as
solution. Perhaps, to complete this work we need automate humans
process described in ~\cite{10.7717/peerj-cs.156}.

The last approach is based on to trust on commercial solutions, such
as we trust on search engines to make searches in Internet (black
box).  In Damegender we can download json files from main commercial
Application Programming Interfaces (API) solutions (genderapi,
genderize, namsor, nameapi, ...). One user can build proprietary
datasets on this way using an average weighted by the precision or
accuracy of each Application Programming Interface measured with
Damegender with an open dataset as base of truth.

\section{Machine Learning}

We have developed a script infofeatures.py with our datasets to
analyze data about features. The datasets used in this experiment was
retrieved from official datasets from national statistical
institutions in Spain, Uruguay, United Kingdom, USA. The features used
are: first letter, last letter, a, b, c, d, e, f, g, h, i, j, k, l, m,
n, o, p, q, r, s, t, u, v, w, x, y, z, vocals, consonants, first
letter, first letter vocal, last letter vocal, last letter consonant,
last letter a. The choosing of features was verified with Principal
Component Analysis.

Take a look to the most informative features results with the
different datasets:
\begin{longtable}[]{@{}llllllll@{}}
  \toprule
Dataset & Char A & Last Char A & Last Char O & Last Char Consonant & Last Char Vocal & First Char Consonant & First Char Vocal & \tabularnewline
  \midrule
  \endhead
 Uruguay (F) \footnote{F is for females and M is for males} &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177 & \tabularnewline
 Uruguay (M) \footnotemark[3] &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228 & \tabularnewline
 Spain (F) \footnotemark[3]  &    0.922 &         0.588 &          0.03 &                 0.271 &             0.728 &                  0.772 &              0.228 & \tabularnewline
 Spain (M) \footnotemark[3]  &    0.818 &          0.03 &         0.268 &                 0.569 &              0.43 &                  0.763 &              0.236 & \tabularnewline
 UK (F) \footnotemark[3]     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235 & \tabularnewline
 UK (M) \footnotemark[3]     &    0.716 &         0.036 &         0.039 &                  0.78 &             0.218 &                  0.799 &                0.2 & \tabularnewline
 USA (F) \footnotemark[3]    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177 & \tabularnewline
 USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159 & \tabularnewline
\bottomrule
\caption{Informative Features in Different Countries}
\label{table:InfoFeatures}
\end{longtable}

The countries where the main language is spanish (Uruguay + Spain) and
english (USA + United Kingdom + Australia) where is having very
similar variation with the features chosen between males and females
with these datasets (remember that this datasets are being extracted
from official statistics provided by the states). In Canada, a country
french centric has different rules with this features.

The letter a is varying 0.2 from males to females in (USA and Uruguay)
and 0.1 from males to females (United Kingdom, Australia and
Spain). The last letter a is varying 0.5 from males to females in
(Australia, Spain) around 0.4 in (USA, United Kingdom) and 0.2 in
Uruguay. The last letter o from females to males is varying 0.2 in
(Spain, Australia) and is equal in (Uruguay, USA, United Kingdom). For
the last letter consonant all countries is giving the result that is
for males, with results from 0.2 to 0.5: Uruguay and USA (0.5), United
Kingdom (0.4), Australia and Spain (0.3). So last letter vocal is
reverse tha last letter consonant. First letter consonant or first
letter vocal is a non significative feature due to so similar results
in english and spanish.

The success with the different algorithms is showed in the next table:

\begin{longtable}[]{@{}lllll@{}}
  \toprule
  Machine Learning Algorithm & Accuracy & Precision & F1score & Recall \tabularnewline
  \midrule
  \endhead

 Support Vector Machines             &    0.879 &     0.972 &   0.972 &    1.0 & \tabularnewline
 Random Forest                       &    0.862 &     0.902 &   0.902 &    1.0 & \tabularnewline
 NLTK (Bayes)                        &    0.862 &     0.902 &   0.902 &    1.0 & \tabularnewline
 Multinomial Navie Bayes             &    0.782 &     0.791 &   0.791 &    1.0 & \tabularnewline
 Tree                                &    0.764 &     0.821 &   0.796 &    1.0 & \tabularnewline
 Stochastic Gradient Distribution    &    0.709 &     0.943 &   0.815 &    1.0 & \tabularnewline
 Gaussian Naive Bayes                &    0.709 &     0.968 &   0.887 &    1.0 & \tabularnewline
 Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0 & \tabularnewline
 AdaBoost                            &    0.698 &     0.965 &   0.815 &    1.0 & \tabularnewline
 Multi Layer Perceptron              &    0.677 &     0.819 &   0.755 &    1.0 & \tabularnewline

\bottomrule
\caption{Machine Learning Algorithms and accuracies measures}
\label{table:MLAccuracies}
\end{longtable}

The results in ~\ref{table:MLAccuracies} shows that using algorithms
as Support Vector Machines or Random Forest against a scientific
dataset created by independent researchers where it is possible to
reach results similar to another commercial solutions about gender
detection tools. Our classifier is binary (only male and female).

We were doing this experiment with NLTK and INE datasets with
accuracies reaching accuracies until 0.745. So it makes sense expect
better results in random datasets augmenting languages and
countries. Due to our solution is not providing arabic or chinesse
alphabets, yet.

So, it's possible infer that Damegender provides a good solution for
nicknames, diminutives, or similar.

\section{State of Art}

\subsection{Comparing Commercial Solutions}

A standard commercial Application Programming Interface (API) can
guess the gender for a single name or a list of names (from a CSV file
or an API call). To express geolocalization you can give surnames, a
country ISO code, or a language. Generally, you can give a probability
and a counter associated to a name and gender in a certain population.

~\cite{10.7717/peerj-cs.156} are proposing a good metrics set to
classify these commercial Application Programming Interfaces (features,
measuring errors and success, ...). The features observed are:
Database size (January 2018), Regular data updates, Handles
unstructured full name strings, Handles surnames, Handles non-Latin
alphabets, Implicit geo-localization, Assignment type, Free
parameters, Open source, Application Programming Interface, Monthly
free requests, Monthly subscription cost (100,000 requests/month)).

In the commercial tools is being used different ways to express
probability (confidence, scale, accuracy, precision, recall, ...).

\subsubsection{Datasets}

In ~\cite{berners2001semantic} a world was envisioned where public
structured data could be interconnected with software agents to
process these data, perhaps only recovering information, but mixed
with distributed artificial intelligence would give a big jump to the
semantic richness to the web.

~\cite{janssen2012benefits} shows serious profits for the states
adopting Open Data in three categories (1) political and social, (2)
economical, (3) operational and technical. So, Open Data is a
breakthrough towards the Semantic Web.

We can find Open Data about names and gender in census of citizens in
states and commercial solutions. Free software packages such as
~\cite{krawetz2006gender} or ~\cite{loper2002nltk} is providing good
datasets about names and gender. So, Damegender incorporates different
lists of names from free software solutions wrote before (Natural
Language ToolKit, Gender Guesser, ...) and from Open Data census
(United Kingdom, USA, Spain, Uruguay, ...).

Wikidata ~\cite{42240} provides a semantic and open database about
Wikipedia allowing retrieve information with Sparql, such as names and
gender.

~\cite{10.7717/peerj-cs.156} describes different ways to build a
dataset on hand looking for names in papers, scientific websites,
wikipedia, biographies, photos, ...)

\subsection*{Free Software}

Before Damegender, only ~\cite{krawetz2006gender} was competing as
Free Software solution with the main commercial Application
Programming Interfaces about gender detection from the name. The best
contribution is the dataset containing 48528 names with a good
classification by countries.

\subsection*{More software about gender}

In some studies, for example, about Twitter or Github, some people can
choose between different ways to detect gender (not only names). So,
we can find gender detection tools from faces in images
(~\cite{ranjan2017hyperface}), from hand written
(~\cite{liwicki2011automatic}), or from speeches
(~\cite{koppel2002automatically}).

\subsection*{Massive Gender Detection}

There are good studies measuring gender in Internet. Some studies are
about gender gap in general (~\cite{robles2014floss},
~\cite{holman2018gender}, ~\cite{dollar1999gender}), Twitter
(~\cite{burger2011discriminating}, ~\cite{mislove2011understanding})
Stackoverflow (~\cite{vasilescu2012gender}), Wikipedia
(~\cite{antin2011gender}, ~\cite{hill2013wikipedia}), Github
(~\cite{vasilescu2015gender}) ...

\section*{Implementation}

We have chosen Python free software tools with a good scientific
impact. Natural Language Toolkit for Natural Language Processing
~\cite{loper2002nltk}. Scikit for Machine Learning
~\cite{pedregosa2011scikit}. Numpy for Numerical Computation
~\cite{van2011numpy}. Matplotlib to visualize results
~\cite{hunter2007matplotlib}. And Perceval ~\cite{duenas2018perceval}
to retrieve information in mailing lists and repositories.

The current result is a Python package contributed to pip to be used
from the console.

The software is using an oriented to objects design with unit testing
for classes and methods using nose and unit testing for Python
commands using Bash.

A summary of current features in the software are:

\begin{itemize}[noitemsep]
\item To deduce the gender about a name in Spanish or English (current
  status) from open census in local.
\item To decide about males and females in strings using different
  machine learning algorithms.
\item To use the main solutions in gender detection (genderize, genderapi,
namsor, nameapi and gender guesser) from a command.
\item To research about why a name is related to males or females with
  statistics. We provide Python commands about study and compare
  gender solutions with: confusion matrix, accuracies, error
  measures. And to decide about features: statistical feature weight,
  principal component analysis, ...
\item To determine gender gap in free software repositories or mailing
  lists (proof of concept)
\end{itemize}

\section*{Conclusions}

The market of gender detection tools is dominated by companies based
on payment services through Application Programming Interfaces with
good results. This market could be modified due to Free Software
tools and Open Data giving more explicative results for the user.

Although machine learning techniques are not new in this field, we
are giving an approach to guess strings not found in a dataset that
currently is classified as unknown and the humans trend to think in
gender terms many strings calling it as nicknames or diminutives.

These previous advances in computer science could be giving support to
study the gender gap in repositories and mailing lists. Another future
work is to create a free and universal dataset with support for all
languages and cultures.

\end{document}

%\bibliographystyle{alpha} 
%% \bibliography{satbibtex}

%% \bibitem[Com79]{Comer-btree}
%% D.~Comer.
%% \newblock The ubiquitous b-tree.
%% \newblock {\em Computing Surveys}, 11(2):121--137, June 1979.

%% \bibitem[Knu73]{Knuth-vol3}
%% D.~E. Knuth.
%% \newblock {\em The Art of Computer Programming -- Volume 3 / Sorting and
%%   Searching}.
%% \newblock Addison-Wesley, 1973.

%% \bibliographystyle{alpha} 
%% \bibliography{bibtex}
%inline the .bbl file directly for mailing to authors.

### 2020-05-14 paper2sattose.tex
Removing everything that is not content

@@
 
-\title{Damegender: Writing and Comparing Gender Detection Tools}
+\title{Damegender: \\ Writing and Comparing Gender Detection Tools}
 
 \author{
-David Arroyo Menéndez \\ GSYC \\ Universidad Rey Juan Carlos I \\ d.arroyome@alumnos.urjc.es
+David Arroyo Men\'endez \\ GSyC \\ Universidad Rey Juan Carlos \\ d.arroyome@alumnos.urjc.es
 \and
-Jesús González Barahona \\ GSYC \\ Universidad Rey Juan Carlos I \\ jgb@gsyc.es
+Jesus Gonzalez-Barahona \\ GSyC \\ Universidad Rey Juan Carlos \\ jgb@gsyc.es
 }
@@
 papers, etc. To achieve gender equality and empower all women and
-girls is a goal in sustanaible development in United Nations, so to
+girls is a goal in sustainable development in United Nations, so to
 measure the gender gap is a previous step to find solutions to reduce
@@
 guess gender from a name. This kind of software has the database
-based on propietary databases and the software is not free, so some
+based on proprietary databases and the software is not free, so some
 scientific works are difficult to reproduce.
@@
 offering a solution with a free license and open data names from
-official census useful to replace, use and/or compare these apis with
+official census useful to replace, use and/or compare these APIs with
 very good results. This tool provides Machine Learning to predict
-strings, that's useful to guess diminutives or nicknames.
+strings, that is useful to guess diminutives or nicknames.
 \end{abstract}
@@
 There are different ways to detect gender from a person name and
-perhaps a surname: census, wikipedia, self-references in trust
-websites, ... The most common way to detect gender from a name is the
+perhaps a surname: census, Wikipedia, self-references in trust
+websites, \ldots The most common way to detect gender from a name is the
 Application Programming Interfaces with a good popularity, for
-example, genderapi, namsor, genderize, ...
+example, genderapi, namsor, genderize, \ldots
 
@@
 \item Treatment with names without census, for example, nicknames,
-  diminutives, ...
+  diminutives, \ldots
 \item Massive gender detection from Internet, for example, mailing
-  lists, software repositories, ...
+  lists, software repositories, \ldots
 \end{itemize}
@@
 \item To evaluate quality of different solutions applying metrics
-  suggested by ~\cite{10.7717/peerj-cs.156}
+  suggested by Santamaría and Mihaljević~\cite{10.7717/peerj-cs.156}
 \item To understand the current technology in detail, I have developed
@@
   census dataset.
-\item To do proof-of-concept tests applying Perceval to detect
+\item To do proof-of-concept tests applying Perceval~\cite{duenas2018perceval} to detect
   gender in mailing lists and software repositories.
@@
 
-\section{First Level Heading}
+\section*{Acknowledgements}
 
-First level headings are all flush left, initial caps, bold and in point
-size 12. One line space before the first level heading and $1/2$ line
-space after the first level heading.
+Here the acknowledgements come...
 
-\subsection{Second Level Heading}
-
-Second level headings must be flush left, initial caps, bold and in point
-size 10. One line space before the second level heading and $1/2$ line
-space after the second level heading.
-
-\subsubsection{Third Level Heading}
-
-Third level headings must be flush left, initial caps and bold.
-One line space before the third level heading and $1/2$ line
-space after the third level heading.
-
-\paragraph{Fourth Level Heading}
-
-Fourth level headings must be flush left, initial caps and roman type.
-One line space before the fourth level heading and $1/2$ line
-space after the fourth level heading.
-
-\subsection{Citations In Text}
-
-Citations within the text should indicate the author's last name and
-year\cite{Knuth-vol3}. Reference style\cite{Comer-btree}
-should follow the style that you are used to using, as long as the
-citation style is consistent.
-
-\subsubsection{Footnotes}
-
-Indicate footnotes with a number\footnote{This is a sample footnote} in
-the text. Place the footnotes at the bottom of the page they appear on.
-Precede the footnote with a vertical rule of 2 inches (12 picas).
-
-\subsubsection{Figures}
-
-All artwork must be centered, neat, clean and legible. Do not use pencil
-or hand-drawn artwork. Figure number and caption always appear after the
-the figure. Place one line space before the figure, one line space
-before the figure caption and one line space after the figure caption.
-The figure caption is initial caps and each figure is numbered
-consecutively.
-
-Make sure that the figure caption does not get separated from the
-figure. Leave extra white space at the bottom of the page to avoid
-splitting the figure and figure caption.
-
-Figure \ref{fig1} shows how to include a figure as encapsulated postscript.
-The source of the figure is in file {\tt fig1.eps}.
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[height=4cm]{fig1}
-\caption{Sample EPS figure }
-\label{fig1}
-\end{center}
-\end{figure}
-
-Below is another figure using LaTeX commands.
-
-\begin{figure}[ht]
-\begin{center}
-\setlength{\unitlength}{1pt}
-\footnotesize
-\begin{picture}(160,80)
-        \put(0,0){\framebox(160,80)[]{}}
-        \put(10,35){\framebox(80,40){}}
-        \put(100,20){\framebox(40,20){}}
-        \put(70,10){\framebox(20,10){}}
-        \put(20,5){\framebox(10,5){}}
-\end{picture}
-\caption{Sample Figure Caption}
-\end{center}
-\end{figure}
-
-\subsubsection{Tables}
-
-All tables must be centered, neat, clean and legible. Do not use pencil
-or hand-drawn tables. Table number and title always appear before the
-table.
-
-One line space before the table title, one line space after the table
-title and one line space after the table. The table title must be
-initial caps and each table numbered consecutively.
-
-\begin{table}[ht]
-\begin{center}
-\caption{Sample Table}
-
-\bigskip
-
-\begin{tabular}{|l|l|r|}
-\hline
-A & B & 1\\ \hline
-C & D & 2\\
-E & F & 3\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsubsection{Handling References}
-
-Use a first level heading for the references. References follow the
-acknowledgements.
-
-\subsubsection{Acknowledgements}
-
-Use a third level heading for the acknowledgements. All acknowledgements
-go at the end of the paper.
-
-%\bibliographystyle{alpha} 
-%\bibliography{samplebib}
-%inline the .bbl file directly for mailing to authors.
-
-\begin{thebibliography}
-
-\bibitem[Com79]{Comer-btree}
-D.~Comer.
-\newblock The ubiquitous b-tree.
-\newblock {\em Computing Surveys}, 11(2):121--137, June 1979.
-
-\bibitem[Knu73]{Knuth-vol3}
-Santamaría, Lucía and Mihaljević, Helena
-\newblock {\em Comparison and benchmark of name-to-gender inference services}.
-\newblock Addison-Wesley, 1973.
-
-\end{thebibliography}
+\bibliographystyle{alpha} 
+\bibliography{bibtex}
 

### 2020-05-14 brief.tex
brief.pdf, brief.tex: creation

--- 2020-05-14 sattose.tex
@@
-\documentclass[a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage{twocolceurws}
+%% Submissions for peer-review must enable line-numbering
+%% using the lineno option in the \documentclass command.
+%%
+%% Preprints and camera-ready submissions do not need
+%% line numbers, and should have this option removed.
+%%
+%% Please note that the line numbering option requires
+%% version 1.1 or newer of the wlpeerj.cls file, and
+%% the corresponding author info requires v1.2
+
+%\documentclass[fleqn,10pt]{llncs} % for journal submissions
+\documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions
 
@@
 
-\author{
-David Arroyo Menéndez \\ GSYC \\ Universidad Rey Juan Carlos I \\ d.arroyome@alumnos.urjc.es
-\and
-Jesús González Barahona \\ GSYC \\ Universidad Rey Juan Carlos I \\ jgb@gsyc.es
-}
+\author[1]{David Arroyo Menéndez}
+\affil[1]{Universidad Rey Juan Carlos}
 
-\institution{}
-
-\begin{document}
-\maketitle
-
+\keywords{Gender gap, Gender detection tools, Software repositories}
 \begin{abstract}
@@
 
-\section{Introduction}
+\begin{document}
+
+\flushbottom
+\maketitle
+\thispagestyle{empty}
+
+\section*{1. Introduction}
 
@@
 The problems addressed are:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item Evaluate quality/price with different commercial solutions.
@@
 In this paper, these problems are faced writing a Python solution for:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item To evaluate quality of different solutions applying metrics
@@
 
-In Section 2, we explain the current solutions to the problems. In
-Section 3, we present the results evaluating the current Application
-Programming Interfaces with our software. In Section 4, we discuss
-attempts and problems releasing with a free license a gender detection
-from name program. In Section 5, we discuss how to obtain Open Datasets
-counting names and gender. In Section 6, we describe our
-machine learning solution. In Section 7, we describe general
-implementation details. Finally, in Section 8 we summarize our
-findings, and describe extensions to the work that we are currently
-exploring.
+\section*{2. State of Art}
 
-\section*{3. Application Programming Interfaces}
+\subsection*{Comparing Commercial Solutions}
 
-\subsection*{Market}
+A standard commercial Application Programming Interface (API) can
+guess the gender for a single name or a list of names (from a CSV file
+or an API call). To express geolocalization you can give surnames, a
+country ISO code, or a language. Generally, you can give a probability
+and a counter associated to a name and gender in a certain population.
 
-We have reproduced to ~\cite{10.7717/peerj-cs.156} and updated on
-27/06/2019 and we are showing the results in
-~\ref{table:featuresAndGenderDetectionToolsByName}.
+~\cite{10.7717/peerj-cs.156} are proposing a good metrics set to
+classify these commercial Application Programming Interfaces
+(features, measuring errors and successes, ...). The features observed
+are: Database size (January 2018), Regular data updates, Handles
+unstructured full name strings, Handles surnames, Handles non-Latin
+alphabets, Implicit geo-localization, Assignment type, Free
+parameters, Open source, Application Programming Interface, Monthly
+free requests, Monthly subscription cost (100,000 requests/month)).
 
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-Feature & Gender API & genderguesser & genderize.io & NameAPI & NamSor & Damegender\tabularnewline
-\midrule
-\endhead
-Database size & 431*10^6 & 48.528 & 114*10^6 & 1.428.345 & 4407*10^6  & 197.271\footnote{This cipher is the sum of names in United States of America, United Kingdom, Uruguay and Spain} \tabularnewline
-Regular data updates & yes & no & yes & yes & yes & yes, dev\tabularnewline
-Unstructured strings & yes & no & no & yes & no & yes\tabularnewline
-Handles surnames & yes & no & no & yes & yes & yes\tabularnewline
-Non-Latin alphabets & partially & no & partially & yes & yes & no\tabularnewline
-Geo-localization & yes & no & no & yes & yes & no\tabularnewline
-Exists locale & yes & yes & yes & yes & yes & yes\tabularnewline
-Assingment type & probabilistic & binary & probabilistic & probabilistic & probabilistic & prob\tabularnewline
-Free params & total, prob & gender & total, prob & confidence & scale & total, prob\tabularnewline
-Guessing with ML & no & no & no & no & no & yes\tabularnewline
-Free license & no & yes & no & no & no & yes\tabularnewline
-API & yes & no & yes & yes & yes & future\tabularnewline
-free requests limited & yes (200) & unlimited & yes (1000) & yes & yes & unlimited\tabularnewline
-\bottomrule
-\caption{Features and gender detection tools by name}
-\label{table:featuresAndGenderDetectionToolsByName}
-\end{longtable}
+It's possible different ways to express probability about the
+succesful (confidence, scale, accuracy, precision, recall, ...),
+confusion matrix to understand where it's successes or fails and
+differente errors measures (error coded, error coded without not
+applicable values, error gender bias, not applicable coded).
 
-All solutions have increased the database size from
-~\cite{10.7717/peerj-cs.156}. NameAPI and GenderAPI is reaching more
-features. The tools with a free license have not many features, so for
-now that will not be the trend in many situations. Today, one good
-solution quality and price is Namsor, which provides unlimited names
-through an Application Programming Interface with many features in the
-task to detect gender from the name.
+\subsection*{Datasets}
 
-\subsection*{Reproducing accuracies and confusion matrix}
+We can divide the next options choosing a dataset: (1) a census
+published with a free license (open census way), (2) a dataset done by
+scientist with a paper in a magazine (scientific way), (3) a dataset
+released with a free license in a free software package (free software
+way), (4) a dataset retrieved from commercial Application Programming
+Interfaces (commercial api way).
+
+In Damegender, we are including Open Data census about names and
+gender, such as INE.es, Uruguay, USA and United Kingdom. We want
+datasets provided by the software package to increment the speed
+retrieving data.
+
+From the user final point of view, the value of using Open Data is
+give a good explanation when we are asking about the gender from a
+name (number of males and females using a specific name in a country)
+versus a probability created by the way explained in
+~\cite{10.7717/peerj-cs.156} or similar.
+
+From the scientific point of view, the value of using Open Data is to
+allow that the experiment can be reviewed by peers on an automatic and
+legal way (using proprietary data the reviewer should request it
+separately to make the review).
+
+A second approach is to build the dataset reviewing the names in
+scientific personal sites, Wikipedia,
+... ~\cite{10.7717/peerj-cs.156}. This approach is valid, but it
+consumes many time and efforts, although could be useful if there not
+a legal way to build the dataset.
+
+A third approach is using a dataset from a popular free software
+solution. For instance, Natural Language Tool Kit is providing 8000
+labeled english names. The classification is male or female. The
+problem again is about don't retrieve data with the social science
+quality of National Statistics Institutes. Another example is Gender
+Guesser a good dataset for international names with different
+categories to define the probability. This approach is similar to use
+a dataset released with a paper in a journal, the advantage is to
+understand and add new names with a solid criteria accepted by the
+scientific community.
+
+We are using the census approach as base of truth to distinguish if a
+name is male or female in a geographical area. Generally, a name has a
+strong weight to determine if it's a male or a female on this way, for
+instance, David is registered 365196 times as male and 0 times as
+female in Spain National Institute of Statistics.
+
+Many countries don't provide Open Data census about gender and names,
+but we envisioned build a Dataset about names and gender free and
+universal working from Gender Guesser dataset and Wikidata as
+solution. Perhaps, to complete this work we need automate humans
+process described in ~\cite{10.7717/peerj-cs.156}.
+
+The last approach is based on to trust on commercial solutions, such
+as we trust on search engines to make searches in Internet (black
+box).  In Damegender we can download json files from main commercial
+Application Programming Interfaces (API) solutions (genderapi,
+genderize, namsor, nameapi, ...). One user can build proprietary
+datasets on this way using an average weighted by the precision or
+accuracy of each Application Programming Interface measured with
+Damegender with an open dataset as base of truth.
+
+Some Open Datasets, such Spain (INE.es), or United States of America
+is providing support to surnames and the relationship with the
+ethnicity. For example, in United States of America is giving a
+probability with the race and in Spain INE.es is giving the number of
+people with a surname with a nationality different to the spanish
+nationality. The propietary solutions is allowing infer the gender
+with the surname as another parameter, due to names such as Andrea
+changes the gender depending the nationality: in Italia would be male,
+but in Spain would be female.
+
+\subsection*{Free Software}
+
+Before Damegender, only ~\cite{krawetz2006gender} was competing as
+Free Software solution with the main commercial Application
+Programming Interfaces about gender detection from the name. The best
+contribution is the dataset containing 48528 names with a good
+classification by countries.
+
+\subsection*{More software about gender}
+
+In some studies, for example, about Twitter or Github, some people can
+choose between different ways to detect gender (not only names). So,
+we can find gender detection tools from faces in images
+(~\cite{ranjan2017hyperface}), from hand written
+(~\cite{liwicki2011automatic}), or from speeches
+(~\cite{koppel2002automatically}).
+
+\subsection*{Massive Gender Detection}
+
+There are good studies measuring gender in Internet. Some studies are
+about gender gap in general (~\cite{robles2014floss},
+~\cite{holman2018gender}, ~\cite{dollar1999gender}), Twitter
+(~\cite{burger2011discriminating}, ~\cite{mislove2011understanding})
+Stackoverflow (~\cite{vasilescu2012gender}), Wikipedia
+(~\cite{antin2011gender}, ~\cite{hill2013wikipedia}), Github
+(~\cite{vasilescu2015gender}) ...
+
+\section*{Reproducing accuracies and confusion matrix}
 
@@
 
-\section*{Datasets}
-
-We can divide the next options choosing a dataset: (1) a census
-published with a free license (open census way), (2) a dataset done by
-scientist with a paper in a magazine (scientific way), (3) a dataset
-released with a free license in a free software package (free software
-way), (4) a dataset retrieved from commercial Application Programming
-Interfaces (commercial api way).
-
-\begin{verbatim}
-$ python3 main.py David --total="ine"
-David gender is male
-363559  males for David from INE.es
-0 females for David from INE.es
-\end{verbatim}
-
-In Damegender, we are including Open Data census about names and
-gender, such as INE.es or USA and United Kingdom (births and dies). We
-want datasets provided by the software package to increment the speed
-retrieving data.
-
-From the user final point of view, the value of using Open Data is
-give a good explanation when we are asking about the gender from a
-name (number of males and females using a specific name in a country)
-versus a probability created by the way explained in
-~\cite{10.7717/peerj-cs.156} or similar.
-
-From the scientific point of view, the value of using Open Data is to
-allow that the experiment can be reviewed by peers on an automatic and
-legal way (using proprietary data the reviewer should request it
-separately to make the review).
-
-A second approach is to build the dataset reviewing the names in
-scientific personal sites, Wikipedia,
-... ~\cite{10.7717/peerj-cs.156}. This approach is valid, but it
-consumes many time and efforts, although could be useful if there not
-a legal way to build the dataset.
-
-A third approach is using a dataset from a popular free software
-solution. For instance, Natural Language Tool Kit is providing 8000
-labeled english names. The classification is male or female. The
-problem again is about don't retrieve data with the social science
-quality of National Statistics Institutes. Another example is Gender
-Guesser a good dataset for international names with different
-categories to define the probability. This approach is similar to use
-a dataset released with a paper in a journal, the advantage is to
-understand and add new names with a solid criteria accepted by the
-scientific community.
-
-We are using the census approach as base of truth to distinguish if a
-name is male or female in a geographical area. Generally, a name has a
-strong weight to determine if it's a male or a female on this way, for
-instance, David is registered 365196 times as male and 0 times as
-female in Spain National Institute of Statistics.
-
-Many countries don't provide Open Data census about gender and names,
-but we envisioned build a Dataset about names and gender free and
-universal working from Gender Guesser dataset and Wikidata as
-solution. Perhaps, to complete this work we need automate humans
-process described in ~\cite{10.7717/peerj-cs.156}.
-
-The last approach is based on to trust on commercial solutions, such
-as we trust on search engines to make searches in Internet (black
-box).  In Damegender we can download json files from main commercial
-Application Programming Interfaces (API) solutions (genderapi,
-genderize, namsor, nameapi, ...). One user can build proprietary
-datasets on this way using an average weighted by the precision or
-accuracy of each Application Programming Interface measured with
-Damegender with an open dataset as base of truth.
-
-\section{Machine Learning}
+\section{5. Machine Learning}
 
@@
   \toprule
-Dataset & Char A & Last Char A & Last Char O & Last Char Consonant & Last Char Vocal & First Char Consonant & First Char Vocal & \tabularnewline
+Dataset & char a & last char a & last char o & last char consonant & last char vocal & first char consonant & first char vocal & \tabularnewline
   \midrule
@@
  Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0 & \tabularnewline
- AdaBoost                            &    0.698 &     0.965 &   0.815 &    1.0 & \tabularnewline
- Multi Layer Perceptron              &    0.677 &     0.819 &   0.755 &    1.0 & \tabularnewline
 
@@
 
-\section{State of Art}
-
-\subsection{Comparing Commercial Solutions}
-
-A standard commercial Application Programming Interface (API) can
-guess the gender for a single name or a list of names (from a CSV file
-or an API call). To express geolocalization you can give surnames, a
-country ISO code, or a language. Generally, you can give a probability
-and a counter associated to a name and gender in a certain population.
-
-~\cite{10.7717/peerj-cs.156} are proposing a good metrics set to
-classify these commercial Application Programming Interfaces (features,
-measuring errors and success, ...). The features observed are:
-Database size (January 2018), Regular data updates, Handles
-unstructured full name strings, Handles surnames, Handles non-Latin
-alphabets, Implicit geo-localization, Assignment type, Free
-parameters, Open source, Application Programming Interface, Monthly
-free requests, Monthly subscription cost (100,000 requests/month)).
-
-In the commercial tools is being used different ways to express
-probability (confidence, scale, accuracy, precision, recall, ...).
-
-\subsubsection{Datasets}
-
-In ~\cite{berners2001semantic} a world was envisioned where public
-structured data could be interconnected with software agents to
-process these data, perhaps only recovering information, but mixed
-with distributed artificial intelligence would give a big jump to the
-semantic richness to the web.
-
-~\cite{janssen2012benefits} shows serious profits for the states
-adopting Open Data in three categories (1) political and social, (2)
-economical, (3) operational and technical. So, Open Data is a
-breakthrough towards the Semantic Web.
-
-We can find Open Data about names and gender in census of citizens in
-states and commercial solutions. Free software packages such as
-~\cite{krawetz2006gender} or ~\cite{loper2002nltk} is providing good
-datasets about names and gender. So, Damegender incorporates different
-lists of names from free software solutions wrote before (Natural
-Language ToolKit, Gender Guesser, ...) and from Open Data census
-(United Kingdom, USA, Spain, Uruguay, ...).
-
-Wikidata ~\cite{42240} provides a semantic and open database about
-Wikipedia allowing retrieve information with Sparql, such as names and
-gender.
-
-~\cite{10.7717/peerj-cs.156} describes different ways to build a
-dataset on hand looking for names in papers, scientific websites,
-wikipedia, biographies, photos, ...)
-
-\subsection*{Free Software}
-
-Before Damegender, only ~\cite{krawetz2006gender} was competing as
-Free Software solution with the main commercial Application
-Programming Interfaces about gender detection from the name. The best
-contribution is the dataset containing 48528 names with a good
-classification by countries.
-
-\subsection*{More software about gender}
-
-In some studies, for example, about Twitter or Github, some people can
-choose between different ways to detect gender (not only names). So,
-we can find gender detection tools from faces in images
-(~\cite{ranjan2017hyperface}), from hand written
-(~\cite{liwicki2011automatic}), or from speeches
-(~\cite{koppel2002automatically}).
-
-\subsection*{Massive Gender Detection}
-
-There are good studies measuring gender in Internet. Some studies are
-about gender gap in general (~\cite{robles2014floss},
-~\cite{holman2018gender}, ~\cite{dollar1999gender}), Twitter
-(~\cite{burger2011discriminating}, ~\cite{mislove2011understanding})
-Stackoverflow (~\cite{vasilescu2012gender}), Wikipedia
-(~\cite{antin2011gender}, ~\cite{hill2013wikipedia}), Github
-(~\cite{vasilescu2015gender}) ...
-
-\section*{Implementation}
+\section*{6. Implementation}
 
@@
 
-\section*{Conclusions}
+\section*{7. Conclusions}
 
@@
 
+\bibliography{bibtex}
+
 \end{document}
-
-%\bibliographystyle{alpha} 
-%% \bibliography{satbibtex}
-
-%% \bibitem[Com79]{Comer-btree}
-%% D.~Comer.
-%% \newblock The ubiquitous b-tree.
-%% \newblock {\em Computing Surveys}, 11(2):121--137, June 1979.
-
-%% \bibitem[Knu73]{Knuth-vol3}
-%% D.~E. Knuth.
-%% \newblock {\em The Art of Computer Programming -- Volume 3 / Sorting and
-%%   Searching}.
-%% \newblock Addison-Wesley, 1973.
-
-%% \bibliographystyle{alpha} 
-%% \bibliography{bibtex}
-%inline the .bbl file directly for mailing to authors.

### 2020-05-14 paper2sattose.tex
Tables fixed. Starting with the text

@@
 \usepackage{twocolceurws}
+\usepackage{xcolor}
 
@@
 
+\newif\ifdraft
+\drafttrue
+%\draftfalse
+\input{macros}
+
 \institution{}
@@
 \begin{abstract}
+
 The variable sex (male or female) is one of most used variables for
@@
 
+\grex{Here we should have a paragraph introducing the problem}
+
 There are different ways to detect gender from a person name and
 perhaps a surname: census, Wikipedia, self-references in trust
-websites, \ldots The most common way to detect gender from a name is the
-Application Programming Interfaces with a good popularity, for
-example, genderapi, namsor, genderize, \ldots
+websites, searches in Google Images.
+Another way to do it is by using one of the existing Application Programming Interfaces (APIs) that exist.
+This paper is about the latter, about their possibilities and limitations.
+Therefore, (i) we evaluate the quality and price of different commercial solutions;
+(ii) we discuss about solutions using free licenses;
+(iii) we investigate what happens with those names without census, for example, nicknames or diminutives; and
+(iv) we elaborate on how massive gender detection from Internet, for example, mailing
+  lists or software repositories, can be done.
 
-The problems addressed are:
+As a result, we contribute with: 
+(i) an evaluation of the quality of different solutions applying metrics
+  suggested by Santamaría and Mihaljević~\cite{10.7717/peerj-cs.156};
+(ii) a tool guessing gender from a name giving support to Spanish and
+  English from the open data census provides by the states built to understand 
+  current technologies in detail;
+(iii) a machine learning solution to strings not found in the
+  census dataset to approach the problem with nicknames and diminutives; and 
+(iv) a proof-of-concept applying Perceval~\cite{duenas2018perceval} to detect
+  gender in mailing lists and software repositories.
+
+\grex{The remainder of this paper is structured as follows...}
+
+\section{State of Art}
+
+\subsection{Comparing Commercial Solutions}
+
+A standard commercial Application Programming Interface (API) can
+guess the gender for a single name or a list of names (from a CSV file
+or an API call). To express geolocalization you can give surnames, a
+country ISO code, or a language. Generally, you can give a probability
+and a counter associated to a name and gender in a certain population.
+
+~\cite{10.7717/peerj-cs.156} are proposing a good metrics set to
+classify these commercial Application Programming Interfaces
+(features, measuring errors and successes, ...). The features observed
+are: Database size (January 2018), Regular data updates, Handles
+unstructured full name strings, Handles surnames, Handles non-Latin
+alphabets, Implicit geo-localization, Assignment type, Free
+parameters, Open source, Application Programming Interface, Monthly
+free requests, Monthly subscription cost (100,000 requests/month)).
+
+It's possible different ways to express probability about the
+succesful (confidence, scale, accuracy, precision, recall, ...),
+confusion matrix to understand where it's successes or fails and
+differente errors measures (error coded, error coded without not
+applicable values, error gender bias, not applicable coded).
+
+\subsection{Datasets}
+
+We can divide the next options choosing a dataset: (1) a census
+published with a free license (open census way), (2) a dataset done by
+scientist with a paper in a magazine (scientific way), (3) a dataset
+released with a free license in a free software package (free software
+way), (4) a dataset retrieved from commercial Application Programming
+Interfaces (commercial api way).
+
+In Damegender, we are including Open Data census about names and
+gender, such as INE.es, Uruguay, USA and United Kingdom. We want
+datasets provided by the software package to increment the speed
+retrieving data.
+
+From the user final point of view, the value of using Open Data is
+give a good explanation when we are asking about the gender from a
+name (number of males and females using a specific name in a country)
+versus a probability created by the way explained in
+~\cite{10.7717/peerj-cs.156} or similar.
+
+From the scientific point of view, the value of using Open Data is to
+allow that the experiment can be reviewed by peers on an automatic and
+legal way (using proprietary data the reviewer should request it
+separately to make the review).
+
+A second approach is to build the dataset reviewing the names in
+scientific personal sites, Wikipedia,
+... ~\cite{10.7717/peerj-cs.156}. This approach is valid, but it
+consumes many time and efforts, although could be useful if there not
+a legal way to build the dataset.
+
+A third approach is using a dataset from a popular free software
+solution. For instance, Natural Language Tool Kit is providing 8000
+labeled english names. The classification is male or female. The
+problem again is about don't retrieve data with the social science
+quality of National Statistics Institutes. Another example is Gender
+Guesser a good dataset for international names with different
+categories to define the probability. This approach is similar to use
+a dataset released with a paper in a journal, the advantage is to
+understand and add new names with a solid criteria accepted by the
+scientific community.
+
+We are using the census approach as base of truth to distinguish if a
+name is male or female in a geographical area. Generally, a name has a
+strong weight to determine if it's a male or a female on this way, for
+instance, David is registered 365196 times as male and 0 times as
+female in Spain National Institute of Statistics.
+
+Many countries don't provide Open Data census about gender and names,
+but we envisioned build a Dataset about names and gender free and
+universal working from Gender Guesser dataset and Wikidata as
+solution. Perhaps, to complete this work we need automate humans
+process described in ~\cite{10.7717/peerj-cs.156}.
+
+The last approach is based on to trust on commercial solutions, such
+as we trust on search engines to make searches in Internet (black
+box).  In Damegender we can download json files from main commercial
+Application Programming Interfaces (API) solutions (genderapi,
+genderize, namsor, nameapi, ...). One user can build proprietary
+datasets on this way using an average weighted by the precision or
+accuracy of each Application Programming Interface measured with
+Damegender with an open dataset as base of truth.
+
+Some Open Datasets, such Spain (INE.es), or United States of America
+is providing support to surnames and the relationship with the
+ethnicity. For example, in United States of America is giving a
+probability with the race and in Spain INE.es is giving the number of
+people with a surname with a nationality different to the spanish
+nationality. The propietary solutions is allowing infer the gender
+with the surname as another parameter, due to names such as Andrea
+changes the gender depending the nationality: in Italia would be male,
+but in Spain would be female.
+
+\subsection{Free Software}
+
+Before Damegender, only ~\cite{krawetz2006gender} was competing as
+Free Software solution with the main commercial Application
+Programming Interfaces about gender detection from the name. The best
+contribution is the dataset containing 48528 names with a good
+classification by countries.
+
+\subsection{More software about gender}
+
+In some studies, for example, about Twitter or Github, some people can
+choose between different ways to detect gender (not only names). So,
+we can find gender detection tools from faces in images
+(~\cite{ranjan2017hyperface}), from hand written
+(~\cite{liwicki2011automatic}), or from speeches
+(~\cite{koppel2002automatically}).
+
+\subsection{Massive Gender Detection}
+
+There are good studies measuring gender in Internet. Some studies are
+about gender gap in general (~\cite{robles2014floss},
+~\cite{holman2018gender}, ~\cite{dollar1999gender}), Twitter
+(~\cite{burger2011discriminating}, ~\cite{mislove2011understanding})
+Stackoverflow (~\cite{vasilescu2012gender}), Wikipedia
+(~\cite{antin2011gender}, ~\cite{hill2013wikipedia}), Github
+(~\cite{vasilescu2015gender}) ...
+
+\section{Reproducing accuracies and confusion matrix}
+
+~\cite{10.7717/peerj-cs.156} explains different ways to determine
+gender from a name by humans and it gives 7000 names applying these
+methods. In this dataset the gender is classified as male, female or
+unknown. We have used this dataset, but only male and female to these
+experiments. We are showing the results in the next table:
+
+%\begin{table*}[]{@{}lllllll@{}}
+\begin{table}
+\footnotesize
+\begin{tabular}[]{lcccc}
+  \hline
+  API & Acc & Prec & F1 & Recall\tabularnewline
+\hline
+Genderapi & 0.969 & 0.972 & 0.964 & 1.0\tabularnewline
+Genderize & 0.927 & 0.976 & 0.966 & 1.0\tabularnewline
+Damegender (SVC)\footnote{SVC is the acronym of Support Vector Classification, the Machine Learning algorithm that Damegender was using with this results} & 0.879 & 0.972 & 0.972 & 1.0\tabularnewline
+Namsor & 0.867 & 0.973 & 0.924 & 1.0\tabularnewline
+Nameapi & 0.830 & 0.974 & 0.905 & 1.0\tabularnewline
+Gender Guesser & 0.774 & 0.985 & 0.872 & 1.0\tabularnewline
+\hline
+\end{tabular}
+\caption{Different accuracies measures}
+\label{table:DifferentAccuraciesMeasures}
+\end{table}
+
+In ~\ref{table:DifferentAccuraciesMeasures} Genderapi and Genderize
+are obtaining the best results, although all solutions is reaching
+results better than 0.8 except Gender Guesser.
+
+\begin{table}
+\footnotesize
+\begin{tabular}[]{lrrrr}
+  \hline
+  APIs          & gender & male & female & undef \tabularnewline
+\hline
+%%\endhead
+Genderapi         & male    & 3589 & 155  &  67 \tabularnewline
+                  & female  & 211  & 1734 &  23 \tabularnewline
+Damegender       & male    & 3663 & 147  &   0 \tabularnewline
+(SVC)\footnotemark[1] & female  & 551  & 1497 &   0 \tabularnewline
+Genderguesser     & male    & 3326 &  139 & 346 \tabularnewline
+                  & female  & 78   & 1686 & 204 \tabularnewline
+Namsor            & male    & 3325 & 139  & 346 \tabularnewline
+                  & female  & 78   & 1686 & 204 \tabularnewline
+Genderize         & male    & 3157 & 242  & 412 \tabularnewline
+                  & female  & 75   & 1742 & 151 \tabularnewline
+Nameapi           & male    & 2627 & 674  & 507 \tabularnewline
+                  & female  & 667  & 1061 & 240 \tabularnewline 
+\hline
+\end{tabular}
+\caption{Confusion matrix tables by APIs}
+\label{table:ConfusionMatrixTables}
+\end{table}
+
+With Damegender has been done a comparison about confusion matrix
+tables depending the API (see ~\ref{table:ConfusionMatrixTables}). If
+we compare these results with the results obtained in
+~\cite{10.7717/peerj-cs.156}, we can understand that the results are
+similar.
+
+Genderapi has similar results, but it is being improved the undefined
+results. In Genderguesser is we are obtaining different results and it
+is strange, because the software has not modified from some years
+ago. In Genderize we are obtaining the same results. In Nameapi the
+guessed results is changing from male to female with more errors. In
+Namsor the results is so similar. Damegender is not guessing undefined
+because we predict with machine learning if the string is not in the
+database.
+
+The most important tools Namsor, Genderapi and Genderize are improving the
+accuracies with respect the previous comparison.
+
+\begin{table*}
+\footnotesize
+\begin{tabular}[]{lrrrr}
+\hline
+API & error & error w/o na & na coded & error gender bias\tabularnewline
+\hline
+%\endhead
+Damegender (SVC)\footnotemark[1] & 0.121 & 0.121 & 0.0 & -0.07\tabularnewline
+GenderApi & 0.167 & 0.167 & 0.0 & -0.167\tabularnewline
+Gender Guesser & 0.225 & 0.027 & 0.204 & 0.003\tabularnewline
+Genderize & 0.276 & 0.261 & 0.0204 & -0.0084 \tabularnewline 
+Namsor & 0.332 & 0.262 & 0.095 & 0.01 \tabularnewline
+Nameapi & 0.361 & 0.267 & 0.129 & 0.001 \tabularnewline
+\hline
+\end{tabular}
+\caption{APIs and Errors}
+\label{table:ApisAndErrors}
+\end{table*}
+
+In the table it is possible to observe a high index of errors in
+Nameapi and Namsor and a low index of errors in GenderApi
+and Damegender.
+
+\section{Machine Learning}
+
+We have developed a script infofeatures.py with our datasets to
+analyze data about features. The datasets used in this experiment was
+retrieved from official datasets from national statistical
+institutions in Spain, Uruguay, United Kingdom, USA. The features used
+are: first letter, last letter, a, b, c, d, e, f, g, h, i, j, k, l, m,
+n, o, p, q, r, s, t, u, v, w, x, y, z, vocals, consonants, first
+letter, first letter vocal, last letter vocal, last letter consonant,
+last letter a. The choosing of features was verified with Principal
+Component Analysis.
+
+Take a look to the most informative features results with the
+different datasets:
+
+\begin{table*}
+\footnotesize
+\begin{tabular}[]{lrrrrrrr}
+  \hline
+Dataset & a & last a & last o & last consonant & last vocal & 1st consonant & 1st vocal  \tabularnewline
+  \hline
+  %\endhead
+ Uruguay (F) \footnote{F is for females and M is for males} &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ Uruguay (M) \footnotemark[3] &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
+ Spain (F) \footnotemark[3]  &    0.922 &         0.588 &          0.03 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
+ Spain (M) \footnotemark[3]  &    0.818 &          0.03 &         0.268 &                 0.569 &              0.43 &                  0.763 &              0.236  \tabularnewline
+ UK (F) \footnotemark[3]     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
+ UK (M) \footnotemark[3]     &    0.716 &         0.036 &         0.039 &                  0.78 &             0.218 &                  0.799 &                0.2  \tabularnewline
+ USA (F) \footnotemark[3]    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159  \tabularnewline
+\hline
+\end{tabular}
+\caption{Informative Features in Different Countries}
+\label{table:InfoFeatures}
+\end{table*}
+
+The countries where the main language is spanish (Uruguay + Spain) and
+english (USA + United Kingdom + Australia) where is having very
+similar variation with the features chosen between males and females
+with these datasets (remember that this datasets are being extracted
+from official statistics provided by the states). In Canada, a country
+french centric has different rules with this features.
+
+The letter a is varying 0.2 from males to females in (USA and Uruguay)
+and 0.1 from males to females (United Kingdom, Australia and
+Spain). The last letter a is varying 0.5 from males to females in
+(Australia, Spain) around 0.4 in (USA, United Kingdom) and 0.2 in
+Uruguay. The last letter o from females to males is varying 0.2 in
+(Spain, Australia) and is equal in (Uruguay, USA, United Kingdom). For
+the last letter consonant all countries is giving the result that is
+for males, with results from 0.2 to 0.5: Uruguay and USA (0.5), United
+Kingdom (0.4), Australia and Spain (0.3). So last letter vocal is
+reverse tha last letter consonant. First letter consonant or first
+letter vocal is a non significative feature due to so similar results
+in english and spanish.
+
+The success with the different algorithms is showed in the next table:
+
+\begin{table}
+\footnotesize
+\begin{tabular}[]{lcccc}
+  \hline
+  ML Algorithm & Acc & Prec & F1 & Recall \tabularnewline
+  \hline
+  %\endhead
+ Support Vector Machines             &    0.879 &     0.972 &   0.972 &    1.0  \tabularnewline
+ Random Forest                       &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
+ NLTK (Bayes)                        &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
+ Multinomial Navie Bayes             &    0.782 &     0.791 &   0.791 &    1.0  \tabularnewline
+ Tree                                &    0.764 &     0.821 &   0.796 &    1.0  \tabularnewline
+ Stoch. Gradient Distrib.    &    0.709 &     0.943 &   0.815 &    1.0  \tabularnewline
+ Gaussian Naive Bayes                &    0.709 &     0.968 &   0.887 &    1.0  \tabularnewline
+ Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0  \tabularnewline
+
+\hline
+\end{tabular}
+\caption{Machine Learning Algorithms and accuracies measures}
+\label{table:MLAccuracies}
+\end{table}
+
+The results in ~\ref{table:MLAccuracies} shows that using algorithms
+as Support Vector Machines or Random Forest against a scientific
+dataset created by independent researchers where it is possible to
+reach results similar to another commercial solutions about gender
+detection tools. Our classifier is binary (only male and female).
+
+We were doing this experiment with NLTK and INE datasets with
+accuracies reaching accuracies until 0.745. So it makes sense expect
+better results in random datasets augmenting languages and
+countries. Due to our solution is not providing arabic or chinesse
+alphabets, yet.
+
+So, it's possible infer that Damegender provides a good solution for
+nicknames, diminutives, or similar.
+
+\section{Implementation}
+
+We have chosen Python free software tools with a good scientific
+impact. Natural Language Toolkit for Natural Language Processing
+~\cite{loper2002nltk}. Scikit for Machine Learning
+~\cite{pedregosa2011scikit}. Numpy for Numerical Computation
+~\cite{van2011numpy}. Matplotlib to visualize results
+~\cite{hunter2007matplotlib}. And Perceval~\cite{duenas2018perceval}
+to retrieve information in mailing lists and repositories.
+
+The current result is a Python package contributed to pip to be used
+from the console.
+
+The software is using an oriented to objects design with unit testing
+for classes and methods using nose and unit testing for Python
+commands using Bash.
+
+A summary of current features in the software are:
+
 \begin{itemize}
-\item Evaluate quality/price with different commercial solutions.
-\item Think about solutions using free licenses.
-\item Treatment with names without census, for example, nicknames,
-  diminutives, \ldots
-\item Massive gender detection from Internet, for example, mailing
-  lists, software repositories, \ldots
+\item To deduce the gender about a name in Spanish or English (current
+  status) from open census in local.
+\item To decide about males and females in strings using different
+  machine learning algorithms.
+\item To use the main solutions in gender detection (genderize, genderapi,
+namsor, nameapi and gender guesser) from a command.
+\item To research about why a name is related to males or females with
+  statistics. We provide Python commands about study and compare
+  gender solutions with: confusion matrix, accuracies, error
+  measures. And to decide about features: statistical feature weight,
+  principal component analysis, ...
+\item To determine gender gap in free software repositories or mailing
+  lists (proof of concept)
 \end{itemize}
 
-In this paper, these problems are faced writing a Python solution for:
+\section*{7. Conclusions}
 
-\begin{itemize}
-\item To evaluate quality of different solutions applying metrics
-  suggested by Santamaría and Mihaljević~\cite{10.7717/peerj-cs.156}
-\item To understand the current technology in detail, I have developed
-  a tool guessing gender from a name giving support to Spanish and
-  English from the open data census provides by the states.
-\item To fix the problem with nicknames and diminutives, we have
-  developed a machine learning solution to strings not found in the
-  census dataset.
-\item To do proof-of-concept tests applying Perceval~\cite{duenas2018perceval} to detect
-  gender in mailing lists and software repositories.
-\end{itemize}
+The market of gender detection tools is dominated by companies based
+on payment services through Application Programming Interfaces with
+good results. This market could be modified due to Free Software
+tools and Open Data giving more explicative results for the user.
+
+Although machine learning techniques are not new in this field, we
+are giving an approach to guess strings not found in a dataset that
+currently is classified as unknown and the humans trend to think in
+gender terms many strings calling it as nicknames or diminutives.
+
+These previous advances in computer science could be giving support to
+study the gender gap in repositories and mailing lists. Another future
+work is to create a free and universal dataset with support for all
+languages and cultures.
 

### 2020-05-15 paper2sattose.tex
Ok, so I have read the paper and see what is to be done.

@@
 
-\title{Damegender: \\ Writing and Comparing Gender Detection Tools}
+\title{[Artifact Presentation] Damegender: \\ Writing and Comparing Gender Detection Tools}
 
@@
 
-\grex{Here we should have a paragraph introducing the problem}
+In recent times, many research investigations have been made on gender
+diversity in the IT domain.
+Examples of these efforts range from participation in Twitter~\cite{burger2011discriminating,mislove2011understanding},
+in Wikipedia ~\cite{antin2011gender,hill2013wikipedia}, in science~\cite{holman2018gender,dollar1999gender}, and more specifically in the software domain in StackOverflow~\cite{vasilescu2012gender}, GitHub~\cite{vasilescu2015gender} and in Free/Libre/Open Source Software development~\cite{robles2014floss}.
+The interest on gender diversity is become more and more relevant, and so does the identification methods that allow to perform comprehensive studies on gender representation in different domains, given the large amounts of data available, in particular from collaborative environments.
 
-There are different ways to detect gender from a person name and
-perhaps a surname: census, Wikipedia, self-references in trust
-websites, searches in Google Images.
-Another way to do it is by using one of the existing Application Programming Interfaces (APIs) that exist.
+There are different ways to detect gender from a person name and perhaps a surname.
+A first, more rudimentary, is based on data extracted from the census, Wikipedia, self-references in trust websites, searches in Google Images, among others.
+Another way to do it is by using one of the existing Application Programming Interfaces (APIs).
 This paper is about the latter, about their possibilities and limitations.
@@
 As a result, we contribute with: 
-(i) an evaluation of the quality of different solutions applying metrics
-  suggested by Santamaría and Mihaljević~\cite{10.7717/peerj-cs.156};
-(ii) a tool guessing gender from a name giving support to Spanish and
+(i) an evaluation of the quality of different solutions applying well-known metrics;
+(ii) a tool, called \texttt{damegender}, guessing gender from a name giving support to Spanish and
   English from the open data census provides by the states built to understand 
@@
   census dataset to approach the problem with nicknames and diminutives; and 
-(iv) a proof-of-concept applying Perceval~\cite{duenas2018perceval} to detect
+(iv) a proof-of-concept of \texttt{damegender} to detect
   gender in mailing lists and software repositories.
@@
 
-A standard commercial Application Programming Interface (API) can
-guess the gender for a single name or a list of names (from a CSV file
-or an API call). To express geolocalization you can give surnames, a
-country ISO code, or a language. Generally, you can give a probability
+A standard commercial Application Programming Interface (API) can guess the gender for a single name or a list of names (from a CSV file or an API call). 
+To express geolocalization you can give surnames, a country ISO code, or a language. Generally, you can give a probability
 and a counter associated to a name and gender in a certain population.
 
-~\cite{10.7717/peerj-cs.156} are proposing a good metrics set to
-classify these commercial Application Programming Interfaces
-(features, measuring errors and successes, ...). The features observed
-are: Database size (January 2018), Regular data updates, Handles
-unstructured full name strings, Handles surnames, Handles non-Latin
-alphabets, Implicit geo-localization, Assignment type, Free
-parameters, Open source, Application Programming Interface, Monthly
-free requests, Monthly subscription cost (100,000 requests/month)).
+Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} propose a set of metrics to classify gender tools (features, measuring errors and successes, among others).
+The features observed are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle
+unstructured full name strings, (iv) if they handles surnames, (v) if they handles non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) the type of assignment, (viii) if free parameters are possible, (ix) if the tool is open source, (x) if they offer an API, (xi) the amount of monthly free requests, and (xii) the monthly subscription cost (calculated for 100,000 requests/month)).
 
-It's possible different ways to express probability about the
-succesful (confidence, scale, accuracy, precision, recall, ...),
-confusion matrix to understand where it's successes or fails and
-differente errors measures (error coded, error coded without not
-applicable values, error gender bias, not applicable coded).
+There are different ways to express the probability of a successful identification (confidence, scale, accuracy, precision, recall, ...), confusion matrix to understand where it is successes or fails and different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded).
 
@@
 
-We can divide the next options choosing a dataset: (1) a census
-published with a free license (open census way), (2) a dataset done by
-scientist with a paper in a magazine (scientific way), (3) a dataset
-released with a free license in a free software package (free software
-way), (4) a dataset retrieved from commercial Application Programming
-Interfaces (commercial api way).
+We can divide the next options choosing a dataset: (1) a census published with a free license (open census way), (2) a dataset done by scientist with a paper in a magazine (scientific way), (3) a dataset released with a free license in a free software package (free software way), (4) a dataset retrieved from commercial APIs (commercial API way).
 
-In Damegender, we are including Open Data census about names and
-gender, such as INE.es, Uruguay, USA and United Kingdom. We want
-datasets provided by the software package to increment the speed
-retrieving data.
+In Damegender, we are including Open Data census about names and gender, such as INE.es, Uruguay, USA and United Kingdom.
+We want datasets provided by the software package to increment the speed retrieving data.
 
-From the user final point of view, the value of using Open Data is
-give a good explanation when we are asking about the gender from a
-name (number of males and females using a specific name in a country)
-versus a probability created by the way explained in
-~\cite{10.7717/peerj-cs.156} or similar.
+From the user final point of view, the value of using Open Data is give a good explanation when we are asking about the gender from a name (number of males and females using a specific name in a country) versus a probability created by the way explained in~\cite{10.7717/peerj-cs.156} or similar.
 
-From the scientific point of view, the value of using Open Data is to
-allow that the experiment can be reviewed by peers on an automatic and
-legal way (using proprietary data the reviewer should request it
-separately to make the review).
+From the scientific point of view, the value of using Open Data is to allow that the experiment can be reviewed by peers on an automatic and legal way (using proprietary data the reviewer should request it separately to make the review).
 
-A second approach is to build the dataset reviewing the names in
-scientific personal sites, Wikipedia,
-... ~\cite{10.7717/peerj-cs.156}. This approach is valid, but it
-consumes many time and efforts, although could be useful if there not
-a legal way to build the dataset.
+A second approach is to build the dataset reviewing the names in scientific personal sites, Wikipedia, ... ~\cite{10.7717/peerj-cs.156}. 
+This approach is valid, but it consumes many time and efforts, although could be useful if there not a legal way to build the dataset.
 
-A third approach is using a dataset from a popular free software
-solution. For instance, Natural Language Tool Kit is providing 8000
-labeled english names. The classification is male or female. The
-problem again is about don't retrieve data with the social science
-quality of National Statistics Institutes. Another example is Gender
-Guesser a good dataset for international names with different
-categories to define the probability. This approach is similar to use
-a dataset released with a paper in a journal, the advantage is to
-understand and add new names with a solid criteria accepted by the
-scientific community.
+A third approach is using a dataset from a popular free software solution. For instance, Natural Language Tool Kit is providing 8000 labeled English names. 
+The classification is male or female. The problem again is about do not retrieve data with the social science quality of National Statistics Institutes. 
+Another example is Gender Guesser a good dataset for international names with different categories to define the probability. 
+This approach is similar to use a dataset released with a paper in a journal, the advantage is to understand and add new names with a solid criteria accepted by the scientific community.
 
-We are using the census approach as base of truth to distinguish if a
-name is male or female in a geographical area. Generally, a name has a
-strong weight to determine if it's a male or a female on this way, for
-instance, David is registered 365196 times as male and 0 times as
-female in Spain National Institute of Statistics.
+We are using the census approach as base of truth to distinguish if a name is male or female in a geographical area.
+Generally, a name has a strong weight to determine if it is a male or a female on this way.
+For instance, David is registered 365196 times as male and 0 times as female in Spain National Institute of Statistics.
 
-Many countries don't provide Open Data census about gender and names,
-but we envisioned build a Dataset about names and gender free and
-universal working from Gender Guesser dataset and Wikidata as
-solution. Perhaps, to complete this work we need automate humans
-process described in ~\cite{10.7717/peerj-cs.156}.
+Many countries don't provide Open Data census about gender and names, but we envisioned build a Dataset about names and gender free and universal working from Gender Guesser dataset and Wikidata as solution.
+Perhaps, to complete this work we need automate humans process described in ~\cite{10.7717/peerj-cs.156}.
 
-The last approach is based on to trust on commercial solutions, such
-as we trust on search engines to make searches in Internet (black
-box).  In Damegender we can download json files from main commercial
-Application Programming Interfaces (API) solutions (genderapi,
-genderize, namsor, nameapi, ...). One user can build proprietary
-datasets on this way using an average weighted by the precision or
-accuracy of each Application Programming Interface measured with
-Damegender with an open dataset as base of truth.
+The last approach is based on to trust on commercial solutions, such as we trust on search engines to make searches in Internet (black box).
+In Damegender we can download JSON files from main commercial Application Programming Interfaces (API) solutions (genderapi,
+genderize, namsor, nameapi, ...).
+One user can build proprietary datasets on this way using an average weighted by the precision or accuracy of each Application Programming Interface measured with Damegender with an open dataset as base of truth.
 
-Some Open Datasets, such Spain (INE.es), or United States of America
-is providing support to surnames and the relationship with the
-ethnicity. For example, in United States of America is giving a
-probability with the race and in Spain INE.es is giving the number of
-people with a surname with a nationality different to the spanish
-nationality. The propietary solutions is allowing infer the gender
-with the surname as another parameter, due to names such as Andrea
-changes the gender depending the nationality: in Italia would be male,
-but in Spain would be female.
+Some Open Datasets, such Spain (INE.es), or United States of America is providing support to surnames and the relationship with the ethnicity. 
+For example, in United States of America is giving a probability with the race and in Spain INE.es is giving the number of
+people with a surname with a nationality different to the Spanish nationality.
+The proprietary solutions is allowing infer the gender with the surname as another parameter, due to names such as Andrea
+changes the gender depending the nationality: in Italy would be male, but in Germany it would be female.
 
@@
 
-Before Damegender, only ~\cite{krawetz2006gender} was competing as
-Free Software solution with the main commercial Application
-Programming Interfaces about gender detection from the name. The best
-contribution is the dataset containing 48528 names with a good
-classification by countries.
+Before Damegender, only ~\cite{krawetz2006gender} was competing as Free Software solution with the main commercial Application
+Programming Interfaces about gender detection from the name. 
+The best contribution is the dataset containing 48528 names with a good classification by countries.
 
@@
 
-In some studies, for example, about Twitter or Github, some people can
-choose between different ways to detect gender (not only names). So,
-we can find gender detection tools from faces in images
-(~\cite{ranjan2017hyperface}), from hand written
-(~\cite{liwicki2011automatic}), or from speeches
-(~\cite{koppel2002automatically}).
-
-\subsection{Massive Gender Detection}
-
-There are good studies measuring gender in Internet. Some studies are
-about gender gap in general (~\cite{robles2014floss},
-~\cite{holman2018gender}, ~\cite{dollar1999gender}), Twitter
-(~\cite{burger2011discriminating}, ~\cite{mislove2011understanding})
-Stackoverflow (~\cite{vasilescu2012gender}), Wikipedia
-(~\cite{antin2011gender}, ~\cite{hill2013wikipedia}), Github
-(~\cite{vasilescu2015gender}) ...
+In some studies, for example, about Twitter or GitHub, some people can choose between different ways to detect gender (not only names). 
+So, we can find gender detection tools from faces in images~\cite{ranjan2017hyperface}, from hand written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 
@@
 
-~\cite{10.7717/peerj-cs.156} explains different ways to determine
-gender from a name by humans and it gives 7000 names applying these
-methods. In this dataset the gender is classified as male, female or
-unknown. We have used this dataset, but only male and female to these
-experiments. We are showing the results in the next table:
+~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and it gives 7000 names applying these methods. 
+In this dataset the gender is classified as male, female or unknown. 
+We have used this dataset, but only male and female to these experiments.
+We are showing the results in the Table~\ref{table:DifferentAccuraciesMeasures}.
 
@@
 
-In ~\ref{table:DifferentAccuraciesMeasures} Genderapi and Genderize
-are obtaining the best results, although all solutions is reaching
-results better than 0.8 except Gender Guesser.
+In Table~\ref{table:DifferentAccuraciesMeasures}, Genderapi and Genderize are obtaining the best results, although all solutions is reaching results better than 0.8 except Gender Guesser.
 
@@
 
-With Damegender has been done a comparison about confusion matrix
-tables depending the API (see ~\ref{table:ConfusionMatrixTables}). If
-we compare these results with the results obtained in
-~\cite{10.7717/peerj-cs.156}, we can understand that the results are
+With Damegender has been done a comparison about confusion matrix tables depending the API (see Table~\ref{table:ConfusionMatrixTables}).
+If we compare these results with the results obtained in~\cite{10.7717/peerj-cs.156}, we can understand that the results are
 similar.
 
-Genderapi has similar results, but it is being improved the undefined
-results. In Genderguesser is we are obtaining different results and it
-is strange, because the software has not modified from some years
-ago. In Genderize we are obtaining the same results. In Nameapi the
-guessed results is changing from male to female with more errors. In
-Namsor the results is so similar. Damegender is not guessing undefined
-because we predict with machine learning if the string is not in the
-database.
+Genderapi has similar results, but it is being improved the undefined results.
+In Genderguesser is we are obtaining different results and it is strange, because the software has not modified from some years
+ago. In Genderize we are obtaining the same results. 
+In Nameapi the guessed results is changing from male to female with more errors. 
+In Namsor the results is similar. 
+Damegender is not guessing undefined because we predict with machine learning if the string is not in the database.
 
-The most important tools Namsor, Genderapi and Genderize are improving the
-accuracies with respect the previous comparison.
+The most important tools Namsor, Genderapi and Genderize are improving the accuracies with respect the previous comparison.
 
@@
 
-In the table it is possible to observe a high index of errors in
-Nameapi and Namsor and a low index of errors in GenderApi
-and Damegender.
+In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in Nameapi and Namsor and a low index of errors in GenderApi and Damegender.
 
@@
 
-We have developed a script infofeatures.py with our datasets to
-analyze data about features. The datasets used in this experiment was
-retrieved from official datasets from national statistical
-institutions in Spain, Uruguay, United Kingdom, USA. The features used
-are: first letter, last letter, a, b, c, d, e, f, g, h, i, j, k, l, m,
-n, o, p, q, r, s, t, u, v, w, x, y, z, vocals, consonants, first
-letter, first letter vocal, last letter vocal, last letter consonant,
-last letter a. The choosing of features was verified with Principal
-Component Analysis.
-
-Take a look to the most informative features results with the
-different datasets:
+We have developed a script infofeatures.py with our datasets to analyze data about features.
+The datasets used in this experiment was retrieved from official datasets from national statistical
+institutions in Spain, Uruguay, United Kingdom, USA.
+The features used are: first letter, last letter, [a-z], vocals, consonants, first letter, first letter vocal, last letter vocal, last letter consonant, last letter a.
+The choosing of features was verified with Principal Component Analysis.
+The most relevant results for the different datasets are offered in Table~\ref{table:InfoFeatures}.
 
@@
 
-The countries where the main language is spanish (Uruguay + Spain) and
-english (USA + United Kingdom + Australia) where is having very
-similar variation with the features chosen between males and females
-with these datasets (remember that this datasets are being extracted
-from official statistics provided by the states). In Canada, a country
-french centric has different rules with this features.
+The countries where the main language is Spanish (Uruguay + Spain) and English (USA + United Kingdom + Australia) where is having very similar variation with the features chosen between males and females with these datasets (remember that these datasets are being extracted from official statistics provided by the states).
+In Canada, a country French centric has different rules with this features.
 
-The letter a is varying 0.2 from males to females in (USA and Uruguay)
-and 0.1 from males to females (United Kingdom, Australia and
-Spain). The last letter a is varying 0.5 from males to females in
-(Australia, Spain) around 0.4 in (USA, United Kingdom) and 0.2 in
-Uruguay. The last letter o from females to males is varying 0.2 in
-(Spain, Australia) and is equal in (Uruguay, USA, United Kingdom). For
-the last letter consonant all countries is giving the result that is
-for males, with results from 0.2 to 0.5: Uruguay and USA (0.5), United
-Kingdom (0.4), Australia and Spain (0.3). So last letter vocal is
-reverse tha last letter consonant. First letter consonant or first
-letter vocal is a non significative feature due to so similar results
-in english and spanish.
-
-The success with the different algorithms is showed in the next table:
+The letter a is varying 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
+The last letter a is varying 0.5 from males to females in (Australia, Spain) around 0.4 in (USA, United Kingdom) and 0.2 in
+Uruguay.
+The last letter o from females to males is varying 0.2 in (Spain, Australia) and is equal in (Uruguay, USA, United Kingdom).
+For the last letter consonant all countries is giving the result that is for males, with results from 0.2 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3). 
+So last letter vocal is reverse than last letter consonant. 
+First letter consonant or first letter vocal is a non significant feature due to so similar results in English and Spanish.
+The relevant metrics with the different algorithms are shown in Table~\ref{table:MLAccuracies}.
 
@@
 
-The results in ~\ref{table:MLAccuracies} shows that using algorithms
-as Support Vector Machines or Random Forest against a scientific
-dataset created by independent researchers where it is possible to
-reach results similar to another commercial solutions about gender
-detection tools. Our classifier is binary (only male and female).
+The results in Table~\ref{table:MLAccuracies} show that using algorithms as Support Vector Machines or Random Forest against a scientific dataset created by independent researchers where it is possible to reach results similar to another commercial solutions about gender detection tools.
+Our classifier is binary (only male and female).
 
-We were doing this experiment with NLTK and INE datasets with
-accuracies reaching accuracies until 0.745. So it makes sense expect
-better results in random datasets augmenting languages and
-countries. Due to our solution is not providing arabic or chinesse
-alphabets, yet.
+We were doing this experiment with NLTK and INE datasets with accuracies reaching accuracies until 0.745.
+So it makes sense expect better results in random datasets augmenting languages and countries.
+However, our solution is not providing Arabic or Chinese alphabets, yet.
 
-So, it's possible infer that Damegender provides a good solution for
-nicknames, diminutives, or similar.
+So, it is possible to infer that Damegender provides a good solution for nicknames, diminutives, or similar.
 
@@
 
-We have chosen Python free software tools with a good scientific
-impact. Natural Language Toolkit for Natural Language Processing
-~\cite{loper2002nltk}. Scikit for Machine Learning
-~\cite{pedregosa2011scikit}. Numpy for Numerical Computation
-~\cite{van2011numpy}. Matplotlib to visualize results
-~\cite{hunter2007matplotlib}. And Perceval~\cite{duenas2018perceval}
-to retrieve information in mailing lists and repositories.
+We have chosen Python free software tools with a good scientific impact.
+Natural Language Toolkit for Natural Language Processing ~\cite{loper2002nltk}.
+Scikit for Machine Learning~\cite{pedregosa2011scikit}. 
+Numpy for Numerical Computation~\cite{van2011numpy}. 
+Matplotlib to visualize results~\cite{hunter2007matplotlib}. 
+And Perceval~\cite{duenas2018perceval} to retrieve information in mailing lists and repositories.
 
-The current result is a Python package contributed to pip to be used
-from the console.
+The current result is a Python package contributed to pip to be used from the console.
 
-The software is using an oriented to objects design with unit testing
-for classes and methods using nose and unit testing for Python
-commands using Bash.
+The software is using an oriented to objects design with unit testing for classes and methods using nose and unit testing for Python commands using Bash.
 
@@
 \begin{itemize}
-\item To deduce the gender about a name in Spanish or English (current
-  status) from open census in local.
-\item To decide about males and females in strings using different
-  machine learning algorithms.
-\item To use the main solutions in gender detection (genderize, genderapi,
-namsor, nameapi and gender guesser) from a command.
-\item To research about why a name is related to males or females with
-  statistics. We provide Python commands about study and compare
-  gender solutions with: confusion matrix, accuracies, error
-  measures. And to decide about features: statistical feature weight,
-  principal component analysis, ...
-\item To determine gender gap in free software repositories or mailing
-  lists (proof of concept)
+\item To deduce the gender about a name in Spanish or English (current   status) from open census in local.
+\item To decide about males and females in strings using different machine learning algorithms.
+\item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
+\item To research about why a name is related to males or females with   statistics.
+We provide Python commands about study and compare gender solutions with: confusion matrix, accuracies, error measures.
+And to decide about features: statistical feature weight, principal component analysis, ...
+\item To determine gender gap in free software repositories or mailing lists (proof of concept)
 \end{itemize}
@@
 
-The market of gender detection tools is dominated by companies based
-on payment services through Application Programming Interfaces with
-good results. This market could be modified due to Free Software
-tools and Open Data giving more explicative results for the user.
+The market of gender detection tools is dominated by companies based on payment services through Application Programming Interfaces with good results. 
+This market could be modified due to Free Software tools and Open Data giving more explicative results for the user.
 
-Although machine learning techniques are not new in this field, we
-are giving an approach to guess strings not found in a dataset that
-currently is classified as unknown and the humans trend to think in
-gender terms many strings calling it as nicknames or diminutives.
+Although machine learning techniques are not new in this field, we are giving an approach to guess strings not found in a dataset that currently is classified as unknown and the humans trend to think in gender terms many strings calling it as nicknames or diminutives.
 
-These previous advances in computer science could be giving support to
-study the gender gap in repositories and mailing lists. Another future
-work is to create a free and universal dataset with support for all
-languages and cultures.
+These previous advances in computer science could be giving support to study the gender gap in repositories and mailing lists.
+Another future work is to create a free and universal dataset with support for all languages and cultures.
 

### 2020-05-15 paper2sattose.tex
Major restructuring, adding a nice table that you have in the damegender github page, etc., etc.

@@
 \usepackage{xcolor}
+\usepackage{url}
+
+\def\infinity{\rotatebox{90}{8}}
 
@@
 \author{
-David Arroyo Men\'endez \\ GSyC \\ Universidad Rey Juan Carlos \\ d.arroyome@alumnos.urjc.es
-\and
-Jesus Gonzalez-Barahona \\ GSyC \\ Universidad Rey Juan Carlos \\ jgb@gsyc.es
+David Arroyo Men\'endez, Jesus Gonzalez-Barahona \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc\}.urjc.es
+%\and
+%Jesus Gonzalez-Barahona \\ GSyC \\ Universidad Rey Juan Carlos \\ jgb@gsyc.es
 }
@@
 
+%A summary of current features of \texttt{damegender} are:
+%
+%\begin{itemize}
+%  \item To deduce the gender about a name in Spanish or English (current status) from open census in local.
+%  \item To decide about males and females in strings using different machine learning algorithms.
+%  \item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
+%  \item To research about why a name is related to males or females with   statistics.
+We provide Python commands about study and compare gender solutions with: confusion matrix, accuracies, error measures.
+And to decide about features: statistical feature weight, principal component analysis, ...
+%  \item To determine gender gap in free software repositories or mailing lists (proof of concept)
+%\end{itemize}
+
 \grex{The remainder of this paper is structured as follows...}
 
-\section{State of Art}
+\section{Damegender}
 
-\subsection{Comparing Commercial Solutions}
+\texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool under a free software license (in particular, the GNU General Public License v3.0). 
+It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLT) for Natural Language Processing~\cite{loper2002nltk}, Scikit for Machine Learning~\cite{pedregosa2011scikit}, Numpy for Numerical Computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
+At its current point it is linked to Perceval~\cite{duenas2018perceval}, a tool specialized in retrieving and gathering data from software repositories, such as git and mailing lists.
 
-A standard commercial Application Programming Interface (API) can guess the gender for a single name or a list of names (from a CSV file or an API call). 
-To express geolocalization you can give surnames, a country ISO code, or a language. Generally, you can give a probability
-and a counter associated to a name and gender in a certain population.
+The software is using an oriented to objects design with unit testing for classes and methods using nose and unit testing for Python commands using Bash. 
+\texttt{damegender} is a Python package that can be installed using PIP (the package installer for Python) from the console.
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} propose a set of metrics to classify gender tools (features, measuring errors and successes, among others).
-The features observed are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle
-unstructured full name strings, (iv) if they handles surnames, (v) if they handles non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) the type of assignment, (viii) if free parameters are possible, (ix) if the tool is open source, (x) if they offer an API, (xi) the amount of monthly free requests, and (xii) the monthly subscription cost (calculated for 100,000 requests/month)).
+The main reason for developing \texttt{damegender} is that there are not many free software tools that help in the identification of gender.
+Before \texttt{damegender}, only \texttt{Gender guesser}\footnote{https://github.com/lead-ratings/gender-guesser} offered a free software solution in this field~\cite{krawetz2006gender}, and the project has not been active for more than three years now.
+The best contribution of \texttt{Gender guesser} is the dataset containing 48,528 names with a good classification by countries\footnote{\url{https://raw.githubusercontent.com/lead-ratings/gender-guesser/master/gender_guesser/data/nam_dict.txt}}.
+
+\subsection{Datasets}
+
+Gender guesser tools apply several methods for estimating the gender from a given name. 
+As a starting point, however, all of them rely on a dataset that contains information on what gender a name usually can be attributed to.
+
+There are several sources to create a databases, being the most common:  
+(1) a census published with a free license (open census way), 
+(2) a dataset released with a free license in a free software package (free software way), 
+(3) a dataset retrieved from commercial APIs (commercial API way), and
+(4) a dataset which is the result of an investigation and that has been released publicly (scientific way).
+
+In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom.
+%We want datasets provided by the software package to increment the speed retrieving data.
+Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
+In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.
+
+Hence, we are using the census approach as base of truth to distinguish if a name is male or female in a geographical area. 
+Generally, a name has a strong weight to determine if it is a male or a female on this way.
+For instance, David is registered 365,196 times as male, but 0 times as female in the data offered by the Spain National Institute of Statistics.
+There are names that heavily depend on the region. 
+For instance, Andrea would be considered a female name in Germany, but a male name in Italy.
+However, many countries do not provide Open Data census about gender and names.
+
+%From the user final point of view, the value of using Open Data is to give a good explanation when we are asking about the gender from a name (number of males and females using a specific name in a country) versus a probability created by the way explained in~\cite{10.7717/peerj-cs.156} or similar.
+
+%From the scientific point of view, the value of using Open Data is to allow that the experiment can be reviewed by peers on an automatic and legal way (using proprietary data the reviewer should request it separately to make the review).
+
+%A second approach is to build the dataset reviewing the names in scientific personal sites, Wikipedia, ... ~\cite{10.7717/peerj-cs.156}. 
+%This approach is valid, but it consumes many time and efforts, although could be useful if there not a legal way to build the dataset.
+
+We have evaluated to include data from the second option (datasets released with a free license). 
+For instance, Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
+Another example is Gender Guesser a good dataset for international names with different categories to define the probability. 
+The problem with these data is that we have observed that they do not have the quality of National Statistics Institutes.
+
+%This approach is similar to use a dataset released with a paper in a journal, the advantage is to understand and add new names with a solid criteria accepted by the scientific community.
+
+The third approach is based on the trust on commercial solutions, in the same way we trust search engines when we make searches in Internet.
+This is because commercial APIs can be seen just a black box, so we do not know where the data comes from and how it has been treated.
+As at this point, commercial APIs offer better results as other solutions, \texttt{damegender} gives the possibility to include data from them.
+Thus, it is possible to download JSON files from the main commercial gender guesser API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
+There are certain uses that are currenly only available in such tools.
+
+% One user can build proprietary datasets on this way using an average weighted by the precision or accuracy of each Application Programming Interface measured with \texttt{damegender} with an open dataset as base of truth.
+
+As a final goal, we envision to build a free dataset with names and gender, that builds on top of \texttt{Gender Guesser} and that can be made available as Wikidata. 
+Perhaps, to complete this work, we need to combine an automated with a manual process as described in~\cite{10.7717/peerj-cs.156}.
+
+\section{Comparison with other tools}
+
+Standard commercial Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
+To express geolocalization the user can also give surnames, a country ISO code, or specify a language.
+Generally, you can give a probability and a counter associated to a name and gender in a certain population.
+
+Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handles surnames, (v) if they handles non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) the type of assignment, (viii) if free parameters are possible, (ix) if the tool is open source, (x) if they offer an API, (xi) the amount of monthly free requests, and (xii) the monthly subscription cost (calculated for 100,000 requests/month)).
+
+\begin{table*}
+\footnotesize
+\begin{tabular}[]{lcccccc}
+\hline
+Service / Tool $->$ & Gender API & gender-guesser & genderize.io & NameAPI & NamSor & damegender\tabularnewline
+\hline
+Database size & 431M & 45K & 114M & 1M & 4G & 57K \tabularnewline
+Regular data updates & Yes & no & No & Yes & Yes & Yes\tabularnewline
+Unstructured full name strings & Yes & No & No & Yes & No & Yes\tabularnewline
+Surnames & Yes & No & No & Yes & Yes & Yes\tabularnewline
+Non-Latin alphabets & Partial & No & Partial & Yes & Yes & No\tabularnewline
+Implicit geo-localization & Yes & No & No & Yes & Yes & No\tabularnewline
+Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
+Assingment type & P & B & P & P & P & P \tabularnewline
+Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
+Prediction & No & No & No & No & No & Yes\tabularnewline
+Free license & No & Yes & No & No & No & Yes\tabularnewline
+REST API & Yes & No & Yes & Yes & Yes & Planned\tabularnewline
+Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
+\hline
+\end{tabular}
+\caption{Comparison of the different features that gender guesser software services and tools offer.
+Assignment type = \{P: Probabilistic; B: Binary\}.
+Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \} }
+\label{table:}
+\end{table*}
 
@@
 
-\subsection{Datasets}
-
-We can divide the next options choosing a dataset: (1) a census published with a free license (open census way), (2) a dataset done by scientist with a paper in a magazine (scientific way), (3) a dataset released with a free license in a free software package (free software way), (4) a dataset retrieved from commercial APIs (commercial API way).
-
-In Damegender, we are including Open Data census about names and gender, such as INE.es, Uruguay, USA and United Kingdom.
-We want datasets provided by the software package to increment the speed retrieving data.
-
-From the user final point of view, the value of using Open Data is give a good explanation when we are asking about the gender from a name (number of males and females using a specific name in a country) versus a probability created by the way explained in~\cite{10.7717/peerj-cs.156} or similar.
-
-From the scientific point of view, the value of using Open Data is to allow that the experiment can be reviewed by peers on an automatic and legal way (using proprietary data the reviewer should request it separately to make the review).
-
-A second approach is to build the dataset reviewing the names in scientific personal sites, Wikipedia, ... ~\cite{10.7717/peerj-cs.156}. 
-This approach is valid, but it consumes many time and efforts, although could be useful if there not a legal way to build the dataset.
-
-A third approach is using a dataset from a popular free software solution. For instance, Natural Language Tool Kit is providing 8000 labeled English names. 
-The classification is male or female. The problem again is about do not retrieve data with the social science quality of National Statistics Institutes. 
-Another example is Gender Guesser a good dataset for international names with different categories to define the probability. 
-This approach is similar to use a dataset released with a paper in a journal, the advantage is to understand and add new names with a solid criteria accepted by the scientific community.
-
-We are using the census approach as base of truth to distinguish if a name is male or female in a geographical area.
-Generally, a name has a strong weight to determine if it is a male or a female on this way.
-For instance, David is registered 365196 times as male and 0 times as female in Spain National Institute of Statistics.
-
-Many countries don't provide Open Data census about gender and names, but we envisioned build a Dataset about names and gender free and universal working from Gender Guesser dataset and Wikidata as solution.
-Perhaps, to complete this work we need automate humans process described in ~\cite{10.7717/peerj-cs.156}.
-
-The last approach is based on to trust on commercial solutions, such as we trust on search engines to make searches in Internet (black box).
-In Damegender we can download JSON files from main commercial Application Programming Interfaces (API) solutions (genderapi,
-genderize, namsor, nameapi, ...).
-One user can build proprietary datasets on this way using an average weighted by the precision or accuracy of each Application Programming Interface measured with Damegender with an open dataset as base of truth.
-
-Some Open Datasets, such Spain (INE.es), or United States of America is providing support to surnames and the relationship with the ethnicity. 
-For example, in United States of America is giving a probability with the race and in Spain INE.es is giving the number of
-people with a surname with a nationality different to the Spanish nationality.
-The proprietary solutions is allowing infer the gender with the surname as another parameter, due to names such as Andrea
-changes the gender depending the nationality: in Italy would be male, but in Germany it would be female.
-
-\subsection{Free Software}
-
-Before Damegender, only ~\cite{krawetz2006gender} was competing as Free Software solution with the main commercial Application
-Programming Interfaces about gender detection from the name. 
-The best contribution is the dataset containing 48528 names with a good classification by countries.
-
-\subsection{More software about gender}
-
-In some studies, for example, about Twitter or GitHub, some people can choose between different ways to detect gender (not only names). 
-So, we can find gender detection tools from faces in images~\cite{ranjan2017hyperface}, from hand written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
-
 \section{Reproducing accuracies and confusion matrix}
 
-~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and it gives 7000 names applying these methods. 
-In this dataset the gender is classified as male, female or unknown. 
-We have used this dataset, but only male and female to these experiments.
-We are showing the results in the Table~\ref{table:DifferentAccuraciesMeasures}.
+Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
+In their dataset, gender is classified as male, female or unknown. 
+We have used this dataset, not considering the unknow variable, for our experiments.
+The results can be observed in the Table~\ref{table:DifferentAccuraciesMeasures}.
 
@@
 
-With Damegender has been done a comparison about confusion matrix tables depending the API (see Table~\ref{table:ConfusionMatrixTables}).
+With \texttt{damegender} has been done a comparison about confusion matrix tables depending the API (see Table~\ref{table:ConfusionMatrixTables}).
 If we compare these results with the results obtained in~\cite{10.7717/peerj-cs.156}, we can understand that the results are
@@
 In Namsor the results is similar. 
-Damegender is not guessing undefined because we predict with machine learning if the string is not in the database.
+\texttt{damegender} is not guessing undefined because we predict with machine learning if the string is not in the database.
 
@@
 
-In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in Nameapi and Namsor and a low index of errors in GenderApi and Damegender.
+In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in Nameapi and Namsor and a low index of errors in GenderApi and \texttt{damegender}.
 
@@
 
-So, it is possible to infer that Damegender provides a good solution for nicknames, diminutives, or similar.
+So, it is possible to infer that \texttt{damegender} provides a good solution for nicknames, diminutives, or similar.
 
-\section{Implementation}
+\section{Limitations and further research}
 
-We have chosen Python free software tools with a good scientific impact.
-Natural Language Toolkit for Natural Language Processing ~\cite{loper2002nltk}.
-Scikit for Machine Learning~\cite{pedregosa2011scikit}. 
-Numpy for Numerical Computation~\cite{van2011numpy}. 
-Matplotlib to visualize results~\cite{hunter2007matplotlib}. 
-And Perceval~\cite{duenas2018perceval} to retrieve information in mailing lists and repositories.
-
-The current result is a Python package contributed to pip to be used from the console.
-
-The software is using an oriented to objects design with unit testing for classes and methods using nose and unit testing for Python commands using Bash.
-
-A summary of current features in the software are:
-
-\begin{itemize}
-\item To deduce the gender about a name in Spanish or English (current   status) from open census in local.
-\item To decide about males and females in strings using different machine learning algorithms.
-\item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
-\item To research about why a name is related to males or females with   statistics.
-We provide Python commands about study and compare gender solutions with: confusion matrix, accuracies, error measures.
-And to decide about features: statistical feature weight, principal component analysis, ...
-\item To determine gender gap in free software repositories or mailing lists (proof of concept)
-\end{itemize}
-
-\section*{7. Conclusions}
+In some studies, for example, about Twitter or GitHub, some people can choose between different ways to detect gender (not only names). 
+So, we can find gender detection tools from faces in images~\cite{ranjan2017hyperface}, from hand written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 

### 2020-05-15 paper2sattose.tex
Reviewed up to machine learning

@@
 
-\section{Comparison with other tools}
-
-Standard commercial Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
-To express geolocalization the user can also give surnames, a country ISO code, or specify a language.
-Generally, you can give a probability and a counter associated to a name and gender in a certain population.
-
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
-The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handles surnames, (v) if they handles non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) the type of assignment, (viii) if free parameters are possible, (ix) if the tool is open source, (x) if they offer an API, (xi) the amount of monthly free requests, and (xii) the monthly subscription cost (calculated for 100,000 requests/month)).
-
-\begin{table*}
+\begin{table*}[ht]
 \footnotesize
@@
 Assignment type = \{P: Probabilistic; B: Binary\}.
-Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \} }
-\label{table:}
+Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}.
+\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.} }
+\label{table:comparison}
 \end{table*}
 
-There are different ways to express the probability of a successful identification (confidence, scale, accuracy, precision, recall, ...), confusion matrix to understand where it is successes or fails and different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded).
+\section{Feature comparison with other tools}
+
+Standard commercial Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
+To express geolocalization the user can also give surnames, a country ISO code, or specify a language.
+Generally, you can give a probability and a counter associated to a name and gender in a certain population.
+
+Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
+
+We have used this comparison framework and have extended it with other tools, including \texttt{damegender} and updated it to today.
+Results can be found in Table~\ref{table:comparison}.
 
@@
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
+There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, recall).
+We se the confusion matrix to understand where the different tools suceed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
+
+Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
 In their dataset, gender is classified as male, female or unknown. 
 We have used this dataset, not considering the unknow variable, for our experiments.
-The results can be observed in the Table~\ref{table:DifferentAccuraciesMeasures}.
+The results can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
+As can be observed, Genderapi and Genderize are obtaining the best results, although all solutions reach results better than 0.8, except for \texttt{Gender Guesser}.
 
@@
 
-In Table~\ref{table:DifferentAccuraciesMeasures}, Genderapi and Genderize are obtaining the best results, although all solutions is reaching results better than 0.8 except Gender Guesser.
+We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
+Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
+The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the accuracies with respect the previous comparison.
+In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
+In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
+For \texttt{Genderize}, we obtain the same results. 
+\texttt{Nameapi}'s results is changing from male to female with more errors. 
+In \texttt{Namsor}, the results are similar. 
+\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.
 
@@
 
-With \texttt{damegender} has been done a comparison about confusion matrix tables depending the API (see Table~\ref{table:ConfusionMatrixTables}).
-If we compare these results with the results obtained in~\cite{10.7717/peerj-cs.156}, we can understand that the results are
-similar.
-
-Genderapi has similar results, but it is being improved the undefined results.
-In Genderguesser is we are obtaining different results and it is strange, because the software has not modified from some years
-ago. In Genderize we are obtaining the same results. 
-In Nameapi the guessed results is changing from male to female with more errors. 
-In Namsor the results is similar. 
-\texttt{damegender} is not guessing undefined because we predict with machine learning if the string is not in the database.
-
-The most important tools Namsor, Genderapi and Genderize are improving the accuracies with respect the previous comparison.
+In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt~Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
 
@@
 \end{table*}
-
-In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in Nameapi and Namsor and a low index of errors in GenderApi and \texttt{damegender}.
 

### 2020-05-15 paper2sattose.tex
Included some comments just in case David has a look at them

@@
 
+\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+
 %\begin{table*}[]{@{}lllllll@{}}
@@
 
-In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt~Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
+In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
+
+\grex{Aqui haria falta indicar por que hacemos esta comparacion y que queremos aprender de ella}
 

### 2020-05-15 paper2sattose.tex
Adding first version of new abstract

@@
 \begin{abstract}
+%Context
+Diversity in software development teams have been identified as one of the main ingredients of a more productive, more healthy software community.
+Thus, the interest of the research community in identifying who is contributing has increased in the last years.
+In the software domain, and although other types of diversity exist, this is especially important for the case of gender.
+% Objective, 
+Given the large amount of publicly available data on the software development process that can be retrieved and analyzed from the Internet (e.g., GitHub, StackOverflow), the importance of having methods and tools that help with large amounts of data would be desirable.
+% Method
+In this paper we present a free software tool, called \texttt{namegender}, which we have conceived to given a name (and optionally some additional data such as geographic information) outputs the gender.
+\texttt{damegender} is based on open databases from official census and uses Machine Learning to guess diminutives or nicknames.
+% Results 
+We have compared \texttt{damegender} with other existing gender guesser APIs, obtaining good results.
+% Conclusions.
 
-The variable sex (male or female) is one of most used variables for
-any study in sociology, but this variable can be hidden in Internet
-communities. The gender detection from a name is an important problem
-in Natural Language Processing to decide if a string labeled as name
-is classified as male or female. An engineer will find useful
-make gender detection from a name retrieving information from social
-networks, mailing lists, instant messaging, software repositories,
-papers, etc. To achieve gender equality and empower all women and
-girls is a goal in sustainable development in United Nations, so to
-measure the gender gap is a previous step to find solutions to reduce
-it.
-
-Nowadays, there are several Application Programming Interfaces to
-guess gender from a name. This kind of software has the database
-based on proprietary databases and the software is not free, so some
-scientific works are difficult to reproduce.
-
-In this paper, we are envisioning how to solve these problems,
-offering a solution with a free license and open data names from
-official census useful to replace, use and/or compare these APIs with
-very good results. This tool provides Machine Learning to predict
-strings, that is useful to guess diminutives or nicknames.
 \end{abstract}

### 2020-05-15 paper2sattose.tex
paper2sattose.tex: davidam first revision

@@
 \author{
-David Arroyo Men\'endez, Jesus Gonzalez-Barahona \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc\}.urjc.es
-%\and
-%Jesus Gonzalez-Barahona \\ GSyC \\ Universidad Rey Juan Carlos \\ jgb@gsyc.es
+David Arroyo Menendez, Jesus Gonzalez-Barahona \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc\}.urjc.es
 }
@@
 % Method
-In this paper we present a free software tool, called \texttt{namegender}, which we have conceived to given a name (and optionally some additional data such as geographic information) outputs the gender.
-\texttt{damegender} is based on open databases from official census and uses Machine Learning to guess diminutives or nicknames.
+In this paper we present a free software tool, called \texttt{damegender}, which we have conceived to given a name outputs the gender and a probability.
+\texttt{damegender} is based on open databases from official census and uses Machine Learning to guess strings not classified as names, such as diminutives or nicknames.
 % Results 
-We have compared \texttt{damegender} with other existing gender guesser APIs, obtaining good results.
+We have compared \texttt{damegender} with other tools, obtaining good results.
 % Conclusions.
@@
 
-In recent times, many research investigations have been made on gender
-diversity in the IT domain.
+In recent times, many research investigations have been made on gender diversity in the IT domain.
+
 Examples of these efforts range from participation in Twitter~\cite{burger2011discriminating,mislove2011understanding},
 in Wikipedia ~\cite{antin2011gender,hill2013wikipedia}, in science~\cite{holman2018gender,dollar1999gender}, and more specifically in the software domain in StackOverflow~\cite{vasilescu2012gender}, GitHub~\cite{vasilescu2015gender} and in Free/Libre/Open Source Software development~\cite{robles2014floss}.
+
 The interest on gender diversity is become more and more relevant, and so does the identification methods that allow to perform comprehensive studies on gender representation in different domains, given the large amounts of data available, in particular from collaborative environments.
@@
 (i) an evaluation of the quality of different solutions applying well-known metrics;
-(ii) a tool, called \texttt{damegender}, guessing gender from a name giving support to Spanish and
-  English from the open data census provides by the states built to understand 
-  current technologies in detail;
-(iii) a machine learning solution to strings not found in the
-  census dataset to approach the problem with nicknames and diminutives; and 
+
+(ii) a tool, called \texttt{damegender}, guessing gender from a name
+  giving support to Spanish and English from the open data census
+  provides by the states built to understand current technologies in
+  detail; this tools has been compared with APIs using an international
+  dataset with good results.
+
+(iii) a machine learning solution to strings not found in the census
+  dataset to approach the problem with nicknames and diminutives; and
+
 (iv) a proof-of-concept of \texttt{damegender} to detect
@@
 
-%A summary of current features of \texttt{damegender} are:
-%
-%\begin{itemize}
-%  \item To deduce the gender about a name in Spanish or English (current status) from open census in local.
-%  \item To decide about males and females in strings using different machine learning algorithms.
-%  \item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
-%  \item To research about why a name is related to males or females with   statistics.
+A summary of current features of \texttt{damegender} are:
+
+\begin{itemize}
+ \item To deduce the gender about a name in Spanish or English (current status) from open census in local.
+ \item To decide about males and females in strings using different machine learning algorithms.
+ \item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
+ \item To research about why a name is related to males or females with statistics.
 We provide Python commands about study and compare gender solutions with: confusion matrix, accuracies, error measures.
 And to decide about features: statistical feature weight, principal component analysis, ...
-%  \item To determine gender gap in free software repositories or mailing lists (proof of concept)
-%\end{itemize}
+ \item To determine gender gap in free software repositories or mailing lists (proof of concept).
+\end{itemize}
 
-\grex{The remainder of this paper is structured as follows...}
+The remainder of this paper is structured as follows:
+
+In Section 2, we explain Damegender as solution to the problem,
+explaining that we are using Open Datasets extracted from the states
+as base of truth. Section 3, presents a feature comparison with other
+tools.  Section 4, reproduces accuracies and confusion matrix with a
+scientific dataset. The section 5 is about how we are using Machine
+Learning in Damegender. The section 6 is about conclusion, limitations
+and futher research.
 
@@
 
-\texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool under a free software license (in particular, the GNU General Public License v3.0). 
-It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLT) for Natural Language Processing~\cite{loper2002nltk}, Scikit for Machine Learning~\cite{pedregosa2011scikit}, Numpy for Numerical Computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
+\texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool under a Free Software license (in particular, the GNU General Public License v3.0). 
+It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLTK) for Natural Language Processing~\cite{loper2002nltk}, Scikit for Machine Learning~\cite{pedregosa2011scikit}, Numpy for Numerical Computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
 At its current point it is linked to Perceval~\cite{duenas2018perceval}, a tool specialized in retrieving and gathering data from software repositories, such as git and mailing lists.

### 2020-05-15 paper2sattose.tex
paper2sattose.pdf, paper2sattose.tex: new revision

@@
 
-In Section 2, we explain Damegender as solution to the problem,
-explaining that we are using Open Datasets extracted from the states
-as base of truth. Section 3, presents a feature comparison with other
-tools.  Section 4, reproduces accuracies and confusion matrix with a
-scientific dataset. The section 5 is about how we are using Machine
-Learning in Damegender. The section 6 is about conclusion, limitations
-and futher research.
+In Section 2, we explain Damegender as solution to the
+problem. Section 3 explains that we are using Open Datasets extracted
+from the states as base of truth. Section 4, presents a feature
+comparison with other tools.  Section 5, reproduces accuracies and
+confusion matrix with a scientific dataset. The section 6 is about how
+we are using Machine Learning in Damegender. The section 7 is about
+conclusion, limitations and futher research.
 
@@
 
-\subsection{Datasets}
+\section{Datasets}
 
@@
 In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom.
-%We want datasets provided by the software package to increment the speed retrieving data.
+The datasets provided by the software package is incrementing the speed retrieving data.
+
 Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
@@
 Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}.
-\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.} }
+}
 \label{table:comparison}
 \end{table*}
+
+\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.}
+\davidam{Si te da tiempo por mi adelante}
 
@@
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offers a framework to classify gender tools.
 The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
@@
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
 In their dataset, gender is classified as male, female or unknown. 
@@
 \grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+\davidam{Esto es un poco largo de explicar.
+Está explicado de una manera abreviada más arriba.
+Si te ves con energías para explicarlo, está explicado en https://easychair.org/publications/preprint/vthL}
 
@@
 
-In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
-
-\grex{Aqui haria falta indicar por que hacemos esta comparacion y que queremos aprender de ella}
+In Table~\ref{table:ApisAndErrors} we can observe the different measures for errors in the APIs.
+The most visible conclusion is an high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
 
@@
 
-Here the acknowledgements come...
+Thanks to:
+Gregorio Robles by the support and suggestions for publish this article in Sattose
+
+Lucía Santamaría and Helena Mihaljevic by the previous work.
+
+Daniel Izquierdo and Laura Arjona for starts this research field in the URJC.
+
+Jesus Gonzalez Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.
 

### 2020-05-15 paper2sattose.tex
Actualizado hasta sección 6

--- 2020-05-15 paper2sattose.tex
@@
 \author{
-David Arroyo Menendez, Jesus Gonzalez-Barahona \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc\}.urjc.es
+David Arroyo Menendez, Jesus Gonzalez-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
 }
@@
 A first, more rudimentary, is based on data extracted from the census, Wikipedia, self-references in trust websites, searches in Google Images, among others.
+So, for instance, in some studies, for example, about Twitter or GitHub, some people have used not only names to detect gender. 
+Thus, we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand -written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
+
 Another way to do it is by using one of the existing Application Programming Interfaces (APIs).
@@
 
-A summary of current features of \texttt{damegender} are:
-
-\begin{itemize}
- \item To deduce the gender about a name in Spanish or English (current status) from open census in local.
- \item To decide about males and females in strings using different machine learning algorithms.
- \item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
- \item To research about why a name is related to males or females with statistics.
-We provide Python commands about study and compare gender solutions with: confusion matrix, accuracies, error measures.
+%A summary of current features of \texttt{damegender} are:
+%
+%\begin{itemize}
+% \item To deduce the gender about a name in Spanish or English (current status) from open census in local.
+% \item To decide about males and females in strings using different machine learning algorithms.
+% \item To use the main solutions in gender detection (genderize, genderapi, namsor, nameapi and gender guesser) from a command.
+% \item To allow research to link names to males or females, based on statistical .
+We provide Python commands to study and compare gender solutions with confusion matrix, accuracies, and error measures.
 And to decide about features: statistical feature weight, principal component analysis, ...
- \item To determine gender gap in free software repositories or mailing lists (proof of concept).
-\end{itemize}
+% \item To determine gender gap in free software repositories or mailing lists (proof of concept).
+%\end{itemize}
 
@@
 
-In Section 2, we explain Damegender as solution to the problem,
-explaining that we are using Open Datasets extracted from the states
-as base of truth. Section 3, presents a feature comparison with other
-tools.  Section 4, reproduces accuracies and confusion matrix with a
-scientific dataset. The section 5 is about how we are using Machine
-Learning in Damegender. The section 6 is about conclusion, limitations
-and futher research.
+In Section 2 we explain \texttt{damegender} as solution to the
+problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
+Section 4 presents a feature comparison with other gender guession tools and services.
+Section 5 reports on accuracies and offers a confusion matrix using a scientific dataset.
+Section 6 is about how we use Machine Learning in \texttt{damegender}.
+Section 7 discusses limitations and futher research, and concludes the paper.
 
@@
 At its current point it is linked to Perceval~\cite{duenas2018perceval}, a tool specialized in retrieving and gathering data from software repositories, such as git and mailing lists.
+\texttt{damegender} is a Python package that can be installed using PIP (the package installer for Python) from the console.
 
-The software is using an oriented to objects design with unit testing for classes and methods using nose and unit testing for Python commands using Bash. 
-\texttt{damegender} is a Python package that can be installed using PIP (the package installer for Python) from the console.
+%The software is using an object-oriented design with unit testing for classes and methods.
 
@@
 In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom.
-%We want datasets provided by the software package to increment the speed retrieving data.
-Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
-In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.
-
-Hence, we are using the census approach as base of truth to distinguish if a name is male or female in a geographical area. 
-Generally, a name has a strong weight to determine if it is a male or a female on this way.
-For instance, David is registered 365,196 times as male, but 0 times as female in the data offered by the Spain National Institute of Statistics.
-There are names that heavily depend on the region. 
-For instance, Andrea would be considered a female name in Germany, but a male name in Italy.
-However, many countries do not provide Open Data census about gender and names.
-
-%From the user final point of view, the value of using Open Data is to give a good explanation when we are asking about the gender from a name (number of males and females using a specific name in a country) versus a probability created by the way explained in~\cite{10.7717/peerj-cs.156} or similar.
-
-%From the scientific point of view, the value of using Open Data is to allow that the experiment can be reviewed by peers on an automatic and legal way (using proprietary data the reviewer should request it separately to make the review).
-
-%A second approach is to build the dataset reviewing the names in scientific personal sites, Wikipedia, ... ~\cite{10.7717/peerj-cs.156}. 
-%This approach is valid, but it consumes many time and efforts, although could be useful if there not a legal way to build the dataset.
-
-We have evaluated to include data from the second option (datasets released with a free license). 
-For instance, Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
-Another example is Gender Guesser a good dataset for international names with different categories to define the probability. 
-The problem with these data is that we have observed that they do not have the quality of National Statistics Institutes.
-
-%This approach is similar to use a dataset released with a paper in a journal, the advantage is to understand and add new names with a solid criteria accepted by the scientific community.
-
-The third approach is based on the trust on commercial solutions, in the same way we trust search engines when we make searches in Internet.
-This is because commercial APIs can be seen just a black box, so we do not know where the data comes from and how it has been treated.
-As at this point, commercial APIs offer better results as other solutions, \texttt{damegender} gives the possibility to include data from them.
-Thus, it is possible to download JSON files from the main commercial gender guesser API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
-There are certain uses that are currenly only available in such tools.
-
-% One user can build proprietary datasets on this way using an average weighted by the precision or accuracy of each Application Programming Interface measured with \texttt{damegender} with an open dataset as base of truth.
-
-As a final goal, we envision to build a free dataset with names and gender, that builds on top of \texttt{Gender Guesser} and that can be made available as Wikidata. 
-Perhaps, to complete this work, we need to combine an automated with a manual process as described in~\cite{10.7717/peerj-cs.156}.
+The datasets provided by the software package is incrementing the speed retrieving data.
 
@@
 Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
+Subscription (100K requests/month)	 & 79 & 0 & 7 & 150 & 80 & 0 \tabularnewline
 \hline
@@
 Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}.
-\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.} }
+The subscription price is given in euro.}
 \label{table:comparison}
 \end{table*}
+
+Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
+In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.
+
+Hence, we are using the census approach as base of truth to distinguish if a name is male or female in a geographical area. 
+Generally, a name has a strong weight to determine if it is a male or a female on this way.
+For instance, David is registered 365,196 times as male, but 0 times as female in the data offered by the Spain National Institute of Statistics.
+There are names that heavily depend on the region. 
+For instance, Andrea would be considered a female name in Germany, but a male name in Italy.
+However, many countries do not provide Open Data census about gender and names.
+
+%From the user final point of view, the value of using Open Data is to give a good explanation when we are asking about the gender from a name (number of males and females using a specific name in a country) versus a probability created by the way explained in~\cite{10.7717/peerj-cs.156} or similar.
+
+%From the scientific point of view, the value of using Open Data is to allow that the experiment can be reviewed by peers on an automatic and legal way (using proprietary data the reviewer should request it separately to make the review).
+
+%A second approach is to build the dataset reviewing the names in scientific personal sites, Wikipedia, ... ~\cite{10.7717/peerj-cs.156}. 
+%This approach is valid, but it consumes many time and efforts, although could be useful if there not a legal way to build the dataset.
+
+We have evaluated to include data from the second option (datasets released with a free license). 
+For instance, Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
+Another example is \texttt{Gender Guesser} a good dataset for international names with different categories to define the probability. 
+The problem with these data is that we have observed that they do not have the quality of National Statistics Institutes.
+
+%This approach is similar to use a dataset released with a paper in a journal, the advantage is to understand and add new names with a solid criteria accepted by the scientific community.
+
+The third approach is based on the trust on commercial solutions, in the same way we trust search engines when we make searches in Internet.
+This is because commercial APIs can be seen just a black box, so we do not know where the data comes from and how it has been treated.
+As at this point, commercial APIs offer better results as other solutions, \texttt{damegender} gives the possibility to include data from them.
+Thus, it is possible to download JSON files from the main commercial gender guesser API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
+There are certain uses that are currenly only available in such tools.
+
+% One user can build proprietary datasets on this way using an average weighted by the precision or accuracy of each Application Programming Interface measured with \texttt{damegender} with an open dataset as base of truth.
+
+As a final goal, we envision to build a free dataset with names and gender, that builds on top of \texttt{Gender Guesser} and that can be made available as Wikidata. 
+Perhaps, to complete this work, we need to combine an automated with a manual process as described in~\cite{10.7717/peerj-cs.156}.
 
@@
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
 The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
@@
 
-Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offer 7,000 names applying these methods. 
 In their dataset, gender is classified as male, female or unknown. 
 We have used this dataset, not considering the unknow variable, for our experiments.
-The results can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
-As can be observed, Genderapi and Genderize are obtaining the best results, although all solutions reach results better than 0.8, except for \texttt{Gender Guesser}.
+The results, using common information retrieval metrics, can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
+Accuracy is the ratio of correctly predicted observation to the total observations.
+It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects.
+So, for a sample with 85 negative and 15 positive values, classifying all values as negative in this case gives a score of accuracy of 0.85.
+In those cases, it is better to report other measures, such as the balanced accuracy (bACC), which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
+Precision is the fraction of relevant instances among the retrieved instances.
+Recall is the fraction of the total amount of relevant instances that were actually retrieved.
+In our case, we have left no name out, so recall is 1 for all tools.
+Precision and recall are sometimes used together in the F1 Score (or f-measure) to provide a single measurement for a system.
+As can be observed, Genderapi and Genderize obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
 
-\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+%\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+%\davidam{Esto es un poco largo de explicar.
+Está explicado de una manera abreviada más arriba.
+Si te ves con energías para explicarlo, está explicado en https://easychair.org/publications/preprint/vthL}
 
 %\begin{table*}[]{@{}lllllll@{}}
-\begin{table}
+\begin{table}[t]
 \footnotesize
@@
 
-\begin{table}
+\begin{table}[t]
 \footnotesize
@@
 
-In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
-
-\grex{Aqui haria falta indicar por que hacemos esta comparacion y que queremos aprender de ella}
+In Table~\ref{table:ApisAndErrors} we observe the different measures for errors in the APIs.
+Error coded defines if the true is different than the guessed one. 
+Error coded without na defines if the true is different than the guessed one, but without undefined results.
+Error gender bias allows to understand if the error is bigger for guessing males than females or viceversa.
+The weighted error defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
+The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
+while \texttt{GenderApi} and \texttt{damegender} have a low index of errors.
 
@@
 \footnotesize
+\center
 \begin{tabular}[]{lrrrr}
@@
 
-We have developed a script infofeatures.py with our datasets to analyze data about features.
-The datasets used in this experiment was retrieved from official datasets from national statistical
-institutions in Spain, Uruguay, United Kingdom, USA.
-The features used are: first letter, last letter, [a-z], vocals, consonants, first letter, first letter vocal, last letter vocal, last letter consonant, last letter a.
-The choosing of features was verified with Principal Component Analysis.
-The most relevant results for the different datasets are offered in Table~\ref{table:InfoFeatures}.
+We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
+To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
+The selection of the features was verified with Principal Component Analysis.
+The datasets used in this experiment were the ones from the Spain National Institute of Statistics and the Natural Language
+ToolKit corpus of (English) names.
+The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.
 
@@
 \footnotesize
-\begin{tabular}[]{lrrrrrrr}
+\center
+\begin{tabular}[]{lccccccc}
   \hline
-Dataset & a & last a & last o & last consonant & last vocal & 1st consonant & 1st vocal  \tabularnewline
+Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
   \hline
   %\endhead
- Uruguay (F) \footnote{F is for females and M is for males} &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
- Uruguay (M) \footnotemark[3] &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
- Spain (F) \footnotemark[3]  &    0.922 &         0.588 &          0.03 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
- Spain (M) \footnotemark[3]  &    0.818 &          0.03 &         0.268 &                 0.569 &              0.43 &                  0.763 &              0.236  \tabularnewline
- UK (F) \footnotemark[3]     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
- UK (M) \footnotemark[3]     &    0.716 &         0.036 &         0.039 &                  0.78 &             0.218 &                  0.799 &                0.2  \tabularnewline
- USA (F) \footnotemark[3]    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
- USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159  \tabularnewline
+ Uruguay (F) &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ Uruguay (M) &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
+ Spain (F)  &    0.922 &         0.588 &          0.030 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
+ Spain (M)  &    0.818 &         0.030 &         0.268 &                 0.569 &              0.430 &                  0.763 &              0.236  \tabularnewline
+ UK (F)     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
+ UK (M)     &    0.716 &         0.036 &         0.039 &                 0.780 &             0.218 &                  0.799 &                0.200  \tabularnewline
+ USA (F)    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ USA (M)    &    0.643 &          0.020 &         0.061 &                 0.765 &             0.234 &                   0.840 &              0.159  \tabularnewline
 \hline
 \end{tabular}
-\caption{Informative Features in Different Countries}
+\caption{Informative Features in Different Countries. F stands for females, and M for males.}
+\vspace{0.3cm}
 \label{table:InfoFeatures}
@@
 
-The countries where the main language is Spanish (Uruguay + Spain) and English (USA + United Kingdom + Australia) where is having very similar variation with the features chosen between males and females with these datasets (remember that these datasets are being extracted from official statistics provided by the states).
-In Canada, a country French centric has different rules with this features.
+As expected, countries that share language offer similar results, i.e., the variation of the chosen featuers between males and females is comparable.
+This is the case for Uruguay and Spain (Spanish, and USA, UK and Australia (English).
+In Canada, a country that has an ample French-speaking community these features show a different trend.
 
-The letter a is varying 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
-The last letter a is varying 0.5 from males to females in (Australia, Spain) around 0.4 in (USA, United Kingdom) and 0.2 in
+For instance, the letter a varies 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
+The last letter a varies 0.5 from males to females in (Australia, Spain), around 0.4 in (USA, United Kingdom) and 0.2 in
 Uruguay.
-The last letter o from females to males is varying 0.2 in (Spain, Australia) and is equal in (Uruguay, USA, United Kingdom).
-For the last letter consonant all countries is giving the result that is for males, with results from 0.2 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3). 
-So last letter vocal is reverse than last letter consonant. 
-First letter consonant or first letter vocal is a non significant feature due to so similar results in English and Spanish.
+The last letter o from females to males varies 0.2 in (Spain, Australia) and is equal in Uruguay, USA, United Kingdom.
+For the last letter consonant all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
+So, last letter vocal is reverse than last letter consonant. 
+First letter consonant or first letter vocal is a non-significant feature due to offering similar results in English and Spanish.
+
 The relevant metrics with the different algorithms are shown in Table~\ref{table:MLAccuracies}.
@@
 \footnotesize
+\center
 \begin{tabular}[]{lcccc}
@@
 
-So, it is possible to infer that \texttt{damegender} provides a good solution for nicknames, diminutives, or similar.
+Based on this experiment, we think that it is possible to infer that \texttt{damegender} could provide a good solution for nicknames, diminutives, or similar.
 
 \section{Limitations and further research}
-
-In some studies, for example, about Twitter or GitHub, some people can choose between different ways to detect gender (not only names). 
-So, we can find gender detection tools from faces in images~\cite{ranjan2017hyperface}, from hand written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 
@@
 
-Here the acknowledgements come...
+Daniel Izquierdo and Laura Arjona for starting this research field at the URJC.
+
+To Jes\'us Gonz\'alez Boticario and the people who work with him at UNED, because of the motivation for the machine learning part.
 

### 2020-05-15 paper2sattose.tex
paper2sattose.tex: add Canada in ML table

--- 2020-05-15 paper2sattose.tex
@@
  USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159  \tabularnewline
+ Canada (F) \footnotemark[3] &    0.659 &         0.189 &         0.005 &                 0.591 &             0.408 &                  0.838 &              0.16  \tabularnewline
+ Canada (M) \footnotemark[3] &    0.752 &          0.22 &         0.025 &                  0.54 &             0.456 &                  0.818 &              0.181  \tabularnewline
+
 \hline

### 2020-05-15 paper2sattose.tex
paper2sattose.tex: start fixes about accents

--- 2020-05-15 paper2sattose.tex
@@
 \documentclass[a4paper]{article}
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
@@
 \author{
-David Arroyo Menendez, Jesus Gonzalez-Barahona \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc\}.urjc.es
+David Arroyo Menéndez, Jesus González-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
 }
@@
 Thanks to:
-Gregorio Robles by the support and suggestions for publish this article in Sattose
 
@@
 
-Jesus Gonzalez Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.
+Jesus González Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.
+
+All people working with Jesus González Barahona and Gregorio Robles.
 

### 2020-05-15 paper2sattose.tex

@@
 
-In Section 2, we explain Damegender as solution to the
-problem. Section 3 explains that we are using Open Datasets extracted
-from the states as base of truth. Section 4, presents a feature
-comparison with other tools.  Section 5, reproduces accuracies and
-confusion matrix with a scientific dataset. The section 6 is about how
-we are using Machine Learning in Damegender. The section 7 is about
-conclusion, limitations and futher research.
+In Section 2 we explain \texttt{damegender} as solution to the
+problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
+Section 4 presents a feature comparison with other gender guession tools and services.
+Section 5 reports on accuracies and offers a confusion matrix using a scientific dataset.
+Section 6 is about how we use Machine Learning in \texttt{damegender}.
+Section 7 discusses limitations and futher research, and concludes the paper.
 
@@
 The datasets provided by the software package is incrementing the speed retrieving data.
+
+\begin{table*}[ht]
+\footnotesize
+\begin{tabular}[]{lcccccc}
+\hline
+Service / Tool $->$ & Gender API & gender-guesser & genderize.io & NameAPI & NamSor & damegender\tabularnewline
+\hline
+Database size & 431M & 45K & 114M & 1M & 4G & 57K \tabularnewline
+Regular data updates & Yes & no & No & Yes & Yes & Yes\tabularnewline
+Unstructured full name strings & Yes & No & No & Yes & No & Yes\tabularnewline
+Surnames & Yes & No & No & Yes & Yes & Yes\tabularnewline
+Non-Latin alphabets & Partial & No & Partial & Yes & Yes & No\tabularnewline
+Implicit geo-localization & Yes & No & No & Yes & Yes & No\tabularnewline
+Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
+Assingment type & P & B & P & P & P & P \tabularnewline
+Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
+Prediction & No & No & No & No & No & Yes\tabularnewline
+Free license & No & Yes & No & No & No & Yes\tabularnewline
+REST API & Yes & No & Yes & Yes & Yes & Planned\tabularnewline
+Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
+Subscription (100K requests/month)	 & 79 & 0 & 7 & 150 & 80 & 0 \tabularnewline
+\hline
+\end{tabular}
+\caption{Comparison of the different features that gender guesser software services and tools offer.
+Assignment type = \{P: Probabilistic; B: Binary\}.
+Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}.
+The subscription price is given in euro.}
+\label{table:comparison}
+\end{table*}
 
@@
 For instance, Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
-Another example is Gender Guesser a good dataset for international names with different categories to define the probability. 
+Another example is \texttt{Gender Guesser} a good dataset for international names with different categories to define the probability. 
 The problem with these data is that we have observed that they do not have the quality of National Statistics Institutes.
@@
 
-\begin{table*}[ht]
-\footnotesize
-\begin{tabular}[]{lcccccc}
-\hline
-Service / Tool $->$ & Gender API & gender-guesser & genderize.io & NameAPI & NamSor & damegender\tabularnewline
-\hline
-Database size & 431M & 45K & 114M & 1M & 4G & 57K \tabularnewline
-Regular data updates & Yes & no & No & Yes & Yes & Yes\tabularnewline
-Unstructured full name strings & Yes & No & No & Yes & No & Yes\tabularnewline
-Surnames & Yes & No & No & Yes & Yes & Yes\tabularnewline
-Non-Latin alphabets & Partial & No & Partial & Yes & Yes & No\tabularnewline
-Implicit geo-localization & Yes & No & No & Yes & Yes & No\tabularnewline
-Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
-Assingment type & P & B & P & P & P & P \tabularnewline
-Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
-Prediction & No & No & No & No & No & Yes\tabularnewline
-Free license & No & Yes & No & No & No & Yes\tabularnewline
-REST API & Yes & No & Yes & Yes & Yes & Planned\tabularnewline
-Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
-\hline
-\end{tabular}
-\caption{Comparison of the different features that gender guesser software services and tools offer.
-Assignment type = \{P: Probabilistic; B: Binary\}.
-Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}.
-}
-\label{table:comparison}
-\end{table*}
-
-\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.}
-\davidam{Si te da tiempo por mi adelante}
-
 \section{Feature comparison with other tools}
@@
 
-Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offers a framework to classify gender tools.
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
 The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
@@
 
-Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offer 7,000 names applying these methods. 
 In their dataset, gender is classified as male, female or unknown. 
 We have used this dataset, not considering the unknow variable, for our experiments.
-The results can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
-As can be observed, Genderapi and Genderize are obtaining the best results, although all solutions reach results better than 0.8, except for \texttt{Gender Guesser}.
+The results, using common information retrieval metrics, can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
+Accuracy is the ratio of correctly predicted observation to the total observations.
+It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects.
+So, for a sample with 85 negative and 15 positive values, classifying all values as negative in this case gives a score of accuracy of 0.85.
+In those cases, it is better to report other measures, such as the balanced accuracy (bACC), which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
+Precision is the fraction of relevant instances among the retrieved instances.
+Recall is the fraction of the total amount of relevant instances that were actually retrieved.
+In our case, we have left no name out, so recall is 1 for all tools.
+Precision and recall are sometimes used together in the F1 Score (or f-measure) to provide a single measurement for a system.
+As can be observed, Genderapi and Genderize obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
 
-\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
-\davidam{Esto es un poco largo de explicar.
+%\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+%\davidam{Esto es un poco largo de explicar.
 Está explicado de una manera abreviada más arriba.
@@
 %\begin{table*}[]{@{}lllllll@{}}
-\begin{table}
+\begin{table}[t]
 \footnotesize
@@
 
-In Table~\ref{table:ApisAndErrors} we can observe the different measures for errors in the APIs.
-The most visible conclusion is an high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.
+In Table~\ref{table:ApisAndErrors} we observe the different measures for errors in the APIs.
+Error coded defines if the true is different than the guessed one. 
+Error coded without na defines if the true is different than the guessed one, but without undefined results.
+Error gender bias allows to understand if the error is bigger for guessing males than females or viceversa.
+The weighted error defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
+The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
+while \texttt{GenderApi} and \texttt{damegender} have a low index of errors.
 
@@
 \footnotesize
-\begin{tabular}[]{lrrrrrrr}
+\center
+\begin{tabular}[]{lccccccc}
   \hline
-Dataset & a & last a & last o & last consonant & last vocal & 1st consonant & 1st vocal  \tabularnewline
+Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
   \hline
@@
  Canada (M) \footnotemark[3] &    0.752 &          0.22 &         0.025 &                  0.54 &             0.456 &                  0.818 &              0.181  \tabularnewline
-
 \hline
 \end{tabular}
-\caption{Informative Features in Different Countries}
+\caption{Informative Features in Different Countries. F stands for females, and M for males.}
+\vspace{0.3cm}
 \label{table:InfoFeatures}

### 2020-05-15 paper2sattose.tex
Trabajando en la parte de ML

@@
 \documentclass[a4paper]{article}
-\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
@@
 \author{
-David Arroyo Menéndez, Jesus González-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
+David Arroyo Menendez, Jesus Gonzalez-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
 }
@@
 Section 4 presents a feature comparison with other gender guession tools and services.
-Section 5 reports on accuracies and offers a confusion matrix using a scientific dataset.
+Section 5 reports on values of accuracy and offers a confusion matrix using a scientific dataset.
 Section 6 is about how we use Machine Learning in \texttt{damegender}.
@@
 
-\section{Reproducing accuracies and confusion matrix}
+\section{Reproducing values of accuracy and confusion matrix}
 
@@
 \end{tabular}
-\caption{Different accuracies measures}
+\caption{Comparison of measures of the quality of the results for the tools under study.}
 \label{table:DifferentAccuraciesMeasures}
@@
 Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
-The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the accuracies with respect the previous comparison.
+The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the values of accuracy with respect the previous comparison.
 In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
@@
 
-We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
-To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
-The selection of the features was verified with Principal Component Analysis.
-The datasets used in this experiment were the ones from the Spain National Institute of Statistics and the Natural Language
-ToolKit corpus of (English) names.
-The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.
-
-\begin{table*}
-\footnotesize
-\center
-\begin{tabular}[]{lccccccc}
-  \hline
-Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
-  \hline
-  %\endhead
- Uruguay (F) \footnote{F is for females and M is for males} &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
- Uruguay (M) \footnotemark[3] &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
- Spain (F) \footnotemark[3]  &    0.922 &         0.588 &          0.03 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
- Spain (M) \footnotemark[3]  &    0.818 &          0.03 &         0.268 &                 0.569 &              0.43 &                  0.763 &              0.236  \tabularnewline
- UK (F) \footnotemark[3]     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
- UK (M) \footnotemark[3]     &    0.716 &         0.036 &         0.039 &                  0.78 &             0.218 &                  0.799 &                0.2  \tabularnewline
- USA (F) \footnotemark[3]    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
- USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159  \tabularnewline
- Canada (F) \footnotemark[3] &    0.659 &         0.189 &         0.005 &                 0.591 &             0.408 &                  0.838 &              0.16  \tabularnewline
- Canada (M) \footnotemark[3] &    0.752 &          0.22 &         0.025 &                  0.54 &             0.456 &                  0.818 &              0.181  \tabularnewline
-\hline
-\end{tabular}
-\caption{Informative Features in Different Countries. F stands for females, and M for males.}
-\vspace{0.3cm}
-\label{table:InfoFeatures}
-\end{table*}
-
-As expected, countries that share language offer similar results, i.e., the variation of the chosen featuers between males and females is comparable.
-This is the case for Uruguay and Spain (Spanish, and USA, UK and Australia (English).
-In Canada, a country that has an ample French-speaking community these features show a different trend.
-
-For instance, the letter a varies 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
-The last letter a varies 0.5 from males to females in (Australia, Spain), around 0.4 in (USA, United Kingdom) and 0.2 in
-Uruguay.
-The last letter o from females to males varies 0.2 in (Spain, Australia) and is equal in Uruguay, USA, United Kingdom.
-For the last letter consonant all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
-So, last letter vocal is reverse than last letter consonant. 
-First letter consonant or first letter vocal is a non-significant feature due to offering similar results in English and Spanish.
-
-The relevant metrics with the different algorithms are shown in Table~\ref{table:MLAccuracies}.
+\subsection{Comparing ML algorithms}
 
@@
 \end{tabular}
-\caption{Machine Learning Algorithms and accuracies measures}
+\caption{Machine Learning Algorithms and accuracy measures}
 \label{table:MLAccuracies}
@@
 
-The results in Table~\ref{table:MLAccuracies} show that using algorithms as Support Vector Machines or Random Forest against a scientific dataset created by independent researchers where it is possible to reach results similar to another commercial solutions about gender detection tools.
+Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
+The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
 Our classifier is binary (only male and female).
 
-We were doing this experiment with NLTK and INE datasets with accuracies reaching accuracies until 0.745.
-So it makes sense expect better results in random datasets augmenting languages and countries.
+\subsection{Experimenting with some features}
+
+We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
+To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
+The selection of the features was verified with Principal Component Analysis.
+The datasets used in this experiment were the ones from the Spain National Institute of Statistics and the Natural Language
+ToolKit corpus of (English) names.
+The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.
+
+\begin{table*}
+\footnotesize
+\center
+\begin{tabular}[]{lccccccc}
+  \hline
+Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
+  \hline
+  %\endhead
+ Uruguay (F) &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ Uruguay (M) &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
+ Spain (F)  &    0.922 &         0.588 &          0.030 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
+ Spain (M)  &    0.818 &         0.030 &         0.268 &                 0.569 &              0.430 &                  0.763 &              0.236  \tabularnewline
+ UK (F)     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
+ UK (M)     &    0.716 &         0.036 &         0.039 &                 0.780 &             0.218 &                  0.799 &                0.200  \tabularnewline
+ USA (F)    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
+ USA (M)    &    0.643 &          0.020 &         0.061 &                 0.765 &             0.234 &                   0.840 &              0.159  \tabularnewline
+\hline
+\end{tabular}
+\caption{Informative Features in Different Countries. F stands for females, and M for males.}
+\vspace{0.3cm}
+\label{table:InfoFeatures}
+\end{table*}
+
+As expected, countries that share language offer similar results, i.e., the variation of the chosen featuers between males and females is comparable.
+This is the case for Uruguay and Spain (Spanish, and USA, UK and Australia (English).
+In Canada, a country that has an ample French-speaking community these features show a different trend.
+
+For instance, the letter a varies 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
+The last letter a varies 0.5 from males to females in (Australia, Spain), around 0.4 in (USA, United Kingdom) and 0.2 in
+Uruguay.
+The last letter o from females to males varies 0.2 in (Spain, Australia) and is equal in Uruguay, USA, United Kingdom.
+For the last letter consonant all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
+So, last letter vocal is reverse than last letter consonant. 
+First letter consonant or first letter vocal is a non-significant feature due to offering similar results in English and Spanish.
+
+We have done this experiment with the NLTK and INE datasets, with the values of accuracy reaching up to 0.745.
+So it makes sense to expect better results in random datasets if we add new languages and countries.
 However, our solution is not providing Arabic or Chinese alphabets, yet.
-
-Based on this experiment, we think that it is possible to infer that \texttt{damegender} could provide a good solution for nicknames, diminutives, or similar.
+The results of this experiment could be used to provide a good solution for nicknames, diminutives, or similar.
 
@@
 
-Thanks to:
+Daniel Izquierdo and Laura Arjona for starting this research field at the URJC.
 
-Lucía Santamaría and Helena Mihaljevic by the previous work.
-
-Daniel Izquierdo and Laura Arjona for starts this research field in the URJC.
-
-Jesus González Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.
-
-All people working with Jesus González Barahona and Gregorio Robles.
+To Jes\'us Gonz\'alez Boticario and the people who work with him at UNED, because of the motivation for the machine learning part.
 

### 2020-05-15 paper2sattose.tex

--- 2020-05-15 paper2sattose.tex
@@
 
+\subsection{Comparing ML algorithms}
+
+\begin{table}
+\footnotesize
+\center
+\begin{tabular}[]{lcccc}
+  \hline
+  ML Algorithm & Acc & Prec & F1 & Recall \tabularnewline
+  \hline
+  %\endhead
+ Support Vector Machines             &    0.879 &     0.972 &   0.972 &    1.0  \tabularnewline
+ Random Forest                       &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
+ NLTK (Bayes)                        &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
+ Multinomial Navie Bayes             &    0.782 &     0.791 &   0.791 &    1.0  \tabularnewline
+ Tree                                &    0.764 &     0.821 &   0.796 &    1.0  \tabularnewline
+ Stoch. Gradient Distrib.    &    0.709 &     0.943 &   0.815 &    1.0  \tabularnewline
+ Gaussian Naive Bayes                &    0.709 &     0.968 &   0.887 &    1.0  \tabularnewline
+ Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0  \tabularnewline
+
+\hline
+\end{tabular}
+\caption{Machine Learning Algorithms and accuracy measures}
+\label{table:MLAccuracies}
+\end{table}
+
+Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
+The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
+Our classifier is binary (only male and female).
+
+\subsection{Experimenting with some features}
+
 We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
@@
 
-The relevant metrics with the different algorithms are shown in Table~\ref{table:MLAccuracies}.
-
-\begin{table}
-\footnotesize
-\center
-\begin{tabular}[]{lcccc}
-  \hline
-  ML Algorithm & Acc & Prec & F1 & Recall \tabularnewline
-  \hline
-  %\endhead
- Support Vector Machines             &    0.879 &     0.972 &   0.972 &    1.0  \tabularnewline
- Random Forest                       &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
- NLTK (Bayes)                        &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
- Multinomial Navie Bayes             &    0.782 &     0.791 &   0.791 &    1.0  \tabularnewline
- Tree                                &    0.764 &     0.821 &   0.796 &    1.0  \tabularnewline
- Stoch. Gradient Distrib.    &    0.709 &     0.943 &   0.815 &    1.0  \tabularnewline
- Gaussian Naive Bayes                &    0.709 &     0.968 &   0.887 &    1.0  \tabularnewline
- Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0  \tabularnewline
-
-\hline
-\end{tabular}
-\caption{Machine Learning Algorithms and accuracies measures}
-\label{table:MLAccuracies}
-\end{table}
-
-The results in Table~\ref{table:MLAccuracies} show that using algorithms as Support Vector Machines or Random Forest against a scientific dataset created by independent researchers where it is possible to reach results similar to another commercial solutions about gender detection tools.
-Our classifier is binary (only male and female).
-
-We were doing this experiment with NLTK and INE datasets with accuracies reaching accuracies until 0.745.
-So it makes sense expect better results in random datasets augmenting languages and countries.
+We have done this experiment with the NLTK and INE datasets, with the values of accuracy reaching up to 0.745.
+So it makes sense to expect better results in random datasets if we add new languages and countries.
 However, our solution is not providing Arabic or Chinese alphabets, yet.
-
-Based on this experiment, we think that it is possible to infer that \texttt{damegender} could provide a good solution for nicknames, diminutives, or similar.
+The results of this experiment could be used to provide a good solution for nicknames, diminutives, or similar.
 

### 2020-05-15 paper2sattose.tex
Beta version

--- 2020-05-15 paper2sattose.tex
@@
 \author{
-David Arroyo Menendez, Jesus Gonzalez-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
+David Arroyo Menéndez, Jesus González-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
 }
@@
 So, for instance, in some studies, for example, about Twitter or GitHub, some people have used not only names to detect gender. 
-Thus, we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand -written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
+Thus, we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand-written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 
@@
 problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
-Section 4 presents a feature comparison with other gender guession tools and services.
+Section 4 presents a feature comparison with other gender guessing tools and services.
 Section 5 reports on values of accuracy and offers a confusion matrix using a scientific dataset.
 Section 6 is about how we use Machine Learning in \texttt{damegender}.
-Section 7 discusses limitations and futher research, and concludes the paper.
+Section 7 discusses limitations and further research, and concludes the paper.
 
@@
 Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
-Assingment type & P & B & P & P & P & P \tabularnewline
+Assignment type & P & B & P & P & P & P \tabularnewline
 Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
@@
 Thus, it is possible to download JSON files from the main commercial gender guesser API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
-There are certain uses that are currenly only available in such tools.
+There are certain uses that are currently only available in such tools.
 
@@
 
-There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, recall).
-We se the confusion matrix to understand where the different tools suceed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
-
-Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offer 7,000 names applying these methods. 
-In their dataset, gender is classified as male, female or unknown. 
-We have used this dataset, not considering the unknow variable, for our experiments.
-The results, using common information retrieval metrics, can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
-Accuracy is the ratio of correctly predicted observation to the total observations.
-It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects.
-So, for a sample with 85 negative and 15 positive values, classifying all values as negative in this case gives a score of accuracy of 0.85.
-In those cases, it is better to report other measures, such as the balanced accuracy (bACC), which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
-Precision is the fraction of relevant instances among the retrieved instances.
-Recall is the fraction of the total amount of relevant instances that were actually retrieved.
-In our case, we have left no name out, so recall is 1 for all tools.
-Precision and recall are sometimes used together in the F1 Score (or f-measure) to provide a single measurement for a system.
-As can be observed, Genderapi and Genderize obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
-
-%\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
-%\davidam{Esto es un poco largo de explicar.
-Está explicado de una manera abreviada más arriba.
-Si te ves con energías para explicarlo, está explicado en https://easychair.org/publications/preprint/vthL}
-
 %\begin{table*}[]{@{}lllllll@{}}
@@
 
-We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
-Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
-The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the values of accuracy with respect the previous comparison.
-In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
-In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
-For \texttt{Genderize}, we obtain the same results. 
-\texttt{Nameapi}'s results is changing from male to female with more errors. 
-In \texttt{Namsor}, the results are similar. 
-\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.
+There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, recall).
+We can see in  the confusion matrix to understand where the different tools succeed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
+
+Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offer 7,000 names applying these methods. 
+In their dataset, gender is classified as male, female or unknown. 
+We have used this dataset, not considering the unknown variable, for our experiments.
+The results, using common information retrieval metrics, can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
+Accuracy is the ratio of correctly predicted observation to the total observations.
+It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects.
+So, for a sample with 85 negative and 15 positive values, classifying all values as negative in this case gives a score of accuracy of 0.85.
+In those cases, it is better to report other measures, such as the balanced accuracy (bACC), which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
+Precision is the fraction of relevant instances among the retrieved instances.
+Recall is the fraction of the total amount of relevant instances that were actually retrieved.
+In our case, we have left no name out, so recall is 1 for all tools.
+Precision and recall are sometimes used together in the F1 Score (or f-measure) to provide a single measurement for a system.
+As can be observed, Genderapi and Genderize obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
+
+%\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
+%\davidam{Esto es un poco largo de explicar.
+Está explicado de una manera abreviada más arriba.
+Si te ves con energías para explicarlo, está explicado en https://easychair.org/publications/preprint/vthL}
 
@@
 
-In Table~\ref{table:ApisAndErrors} we observe the different measures for errors in the APIs.
-Error coded defines if the true is different than the guessed one. 
-Error coded without na defines if the true is different than the guessed one, but without undefined results.
-Error gender bias allows to understand if the error is bigger for guessing males than females or viceversa.
-The weighted error defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
-The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
-while \texttt{GenderApi} and \texttt{damegender} have a low index of errors.
+We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
+Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
+The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the values of accuracy with respect the previous comparison.
+In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
+In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
+For \texttt{Genderize}, we obtain the same results. 
+\texttt{Nameapi}'s results is changing from male to female with more errors. 
+In \texttt{Namsor}, the results are similar. 
+\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.
 
@@
 \caption{APIs and Errors}
+\vspace{0.3cm}
 \label{table:ApisAndErrors}
 \end{table*}
+
+In Table~\ref{table:ApisAndErrors} we observe the different measures for errors in the APIs.
+Error coded defines if the true is different than the guessed one. 
+Error coded without na defines if the true is different than the guessed one, but without undefined results.
+Error gender bias allows to understand if the error is bigger for guessing males than females or viceversa.
+The weighted error defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
+The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
+while \texttt{GenderApi} and \texttt{damegender} have a low index of errors.
 
@@
  USA (M)     &    0.643 &          0.020 &         0.061 &                 0.765 &             0.234 &                   0.840 &              0.159  \tabularnewline
+ Canada (F)  &    0.659 &         0.189 &         0.005 &                 0.591 &             0.408 &                  0.838 &              0.160  \tabularnewline
+ Canada (M)  &    0.752 &          0.220 &         0.025 &                  0.540 &             0.456 &                  0.818 &              0.181  \tabularnewline
+Australia (F)      &    0.922 &         0.588 &         0.033 &                 0.272 &             0.728 &                  0.772 &              0.228 \tabularnewline
+ Australia (M)        &    0.818 &          0.030 &         0.269 &                  0.570 &              0.430 &                  0.763 &              0.237 \tabularnewline
+
 \hline
 \end{tabular}
-\caption{Informative Features in Different Countries. F stands for females, and M for males.}
+\caption{Informative features for different countries. F stands for females, and M for males.}
 \vspace{0.3cm}
@@
 
-As expected, countries that share language offer similar results, i.e., the variation of the chosen featuers between males and females is comparable.
-This is the case for Uruguay and Spain (Spanish, and USA, UK and Australia (English).
+As expected, countries that share language offer similar results, i.e., the variation of the chosen features between males and females is comparable.
+This is the case for Uruguay and Spain (Spanish), and USA and UK and Australia (English).
 In Canada, a country that has an ample French-speaking community these features show a different trend.
 
-For instance, the letter a varies 0.2 from males to females in (USA and Uruguay) and 0.1 from males to females (United Kingdom, Australia and Spain).
-The last letter a varies 0.5 from males to females in (Australia, Spain), around 0.4 in (USA, United Kingdom) and 0.2 in
+For instance, the letter a varies 0.2 from males to females for USA and Uruguay, and 0.1 from males to females for United Kingdom, Australia and Spain.
+The last letter a varies 0.5 from males to females for Australia and Spain, around 0.4 for USA and United Kingdom and 0.2 in
 Uruguay.
@@
 
-The market of gender detection tools is dominated by companies based on payment services through Application Programming Interfaces with good results. 
-This market could be modified due to Free Software tools and Open Data giving more explicative results for the user.
+The market of gender detection tools and services is currently dominated by companies based on payment services through Application Programming Interfaces.
+Without doubt, they offer good results, with high accuracy values.
+However, their inner working cannot be studied (i.e., they work as a black-box for the outsider) and the fees that have to be paid for using their service are sometimes out of the reach of many researchers.
+That is why we propose a new tool, called \texttt{damegender}, with the aim of having open data on gender guessing, which offers more flexibility and where researchers can build on top of it.
+This tool is offered under a free software license and is available on GitHub for download and enhancement.
+As we have shown in this paper, although still incipient, the tool offers good accuracy values based on the use of public databases from government bodies and on the use of machine learning algorithms.
+Nonetheless, we have to note that \texttt{damegender} is still under development, and that it has to be applied to several real repositories to confirm its benefits and address its limitations (such as a small database size of gendered names).
 
-Although machine learning techniques are not new in this field, we are giving an approach to guess strings not found in a dataset that currently is classified as unknown and the humans trend to think in gender terms many strings calling it as nicknames or diminutives.
+In addition, we have shown a glimpse of how several features of the names could be used to guess the gender if we do not have the real name, but nicknames or diminutives.
+This experiment is at this point very preliminary, and we would like to work further on it.
 
-These previous advances in computer science could be giving support to study the gender gap in repositories and mailing lists.
-Another future work is to create a free and universal dataset with support for all languages and cultures.
+All in all, we hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including gender gap) and diversity in the IT, and in particular in the free software community.
+We have therefore many hopes in linking the output of repositories like the ones that can be fetched by Perceval (git, mbox mailing lists, Gerrit, Bugzilla, etc.).
+As a result, we envision a free and universal dataset with support for all languages and cultures.
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 
-Daniel Izquierdo and Laura Arjona for starting this research field at the URJC.
-
-To Jes\'us Gonz\'alez Boticario and the people who work with him at UNED, because of the motivation for the machine learning part.
+We would like to thank: Lucía Santamaría and Helena Mihaljevic by the previous work; Daniel Izquierdo and Laura Arjona for starts this research field in the URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
 

### 2020-05-15 paper2sattose.tex
Beta version2

@@
 
-We would like to thank: Lucía Santamaría and Helena Mihaljevic by the previous work; Daniel Izquierdo and Laura Arjona for starts this research field in the URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
+We would like to thank: Lucía Santamaría and Helena Mihaljevic by the previous work; Daniel Izquierdo and Laura Arjona for starting this research field at URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
 

### 2020-05-15 paper2sattose.tex
paper2sattose.tex: fix datasets in features experiment

--- 2020-05-15 paper2sattose.tex
@@
 The selection of the features was verified with Principal Component Analysis.
-The datasets used in this experiment were the ones from the Spain National Institute of Statistics and the Natural Language
-ToolKit corpus of (English) names.
+The datasets used in this experiment were the ones from the National Institute of Statistics (Spain, Uruguay, United Kingdom, United States of America, Australia and Canada).
 The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.

### 2020-05-15 paper2sattose.tex
Version submitted

@@
 % Conclusions.
-
+We hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including gender gap) and diversity in the IT, and in particular in the free software community.
 \end{abstract}
@@
 In Section 2 we explain \texttt{damegender} as solution to the
-problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
-Section 4 presents a feature comparison with other gender guessing tools and services.
+problem of name-to-gender inference. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
+Section 4 presents a feature comparison with other name-to-gender inference services and tools.
 Section 5 reports on values of accuracy and offers a confusion matrix using a scientific dataset.
@@
 
-Gender guesser tools apply several methods for estimating the gender from a given name. 
+Name-to-gender inference services and tools apply several methods for estimating the gender from a given name. 
 As a starting point, however, all of them rely on a dataset that contains information on what gender a name usually can be attributed to.
@@
 \end{tabular}
-\caption{Comparison of the different features that gender guesser software services and tools offer.
+\caption{Comparison of the different features that name-to-gender inference services and tools offer.
 Assignment type = \{P: Probabilistic; B: Binary\}.
@@
 As at this point, commercial APIs offer better results as other solutions, \texttt{damegender} gives the possibility to include data from them.
-Thus, it is possible to download JSON files from the main commercial gender guesser API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
+Thus, it is possible to download JSON files from the main commercial name-to-gender inference API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
 There are certain uses that are currently only available in such tools.
@@
 
-Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
-The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
-Our classifier is binary (only male and female).
-
-\subsection{Experimenting with some features}
-
-We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
-To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
-The selection of the features was verified with Principal Component Analysis.
-The datasets used in this experiment were the ones from the National Institute of Statistics (Spain, Uruguay, United Kingdom, United States of America, Australia and Canada).
-The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.
-
 \begin{table*}
@@
 
+Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
+The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
+Our classifier is binary (only male and female).
+
+\subsection{Experimenting with some features}
+
+We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
+To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
+The selection of the features was verified with Principal Component Analysis.
+The datasets used in this experiment were the ones from the National Institute of Statistics (Spain, Uruguay, United Kingdom, United States of America, Australia and Canada).
+The most relevant results for the different datasets used are offered in Table~\ref{table:InfoFeatures}.
+
 As expected, countries that share language offer similar results, i.e., the variation of the chosen features between males and females is comparable.
@@
 However, their inner working cannot be studied (i.e., they work as a black-box for the outsider) and the fees that have to be paid for using their service are sometimes out of the reach of many researchers.
-That is why we propose a new tool, called \texttt{damegender}, with the aim of having open data on gender guessing, which offers more flexibility and where researchers can build on top of it.
+That is why we propose a new tool, called \texttt{damegender}, with the aim of having open data on name-to-gender inference, which offers more flexibility and where researchers can build on top of it.
 This tool is offered under a free software license and is available on GitHub for download and enhancement.
@@
 
-We would like to thank: Lucía Santamaría and Helena Mihaljevic by the previous work; Daniel Izquierdo and Laura Arjona for starting this research field at URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
+We would like to thank: Lucía Santamaría and Helena Mihaljevic for the previous work; Daniel Izquierdo and Laura Arjona for starting this research field at URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
 

### 2020-05-16 paper2sattose.tex
Reviewed up to section 4

@@
 %Context
-Diversity in software development teams have been identified as one of the main ingredients of a more productive, more healthy software community.
-Thus, the interest of the research community in identifying who is contributing has increased in the last years.
-In the software domain, and although other types of diversity exist, this is especially important for the case of gender.
+Diversity in software development teams has been identified as one of the main ingredients of a more productive, more healthy software community.
+Thus, the interest of the research community in identifying who is contributing to a project has increased in the last years.
+In the software domain, and although other types of diversity exist, this is particularly important for the case of gender.
 % Objective, 
-Given the large amount of publicly available data on the software development process that can be retrieved and analyzed from the Internet (e.g., GitHub, StackOverflow), the importance of having methods and tools that help with large amounts of data would be desirable.
-% Method
-In this paper we present a free software tool, called \texttt{damegender}, which we have conceived to given a name outputs the gender and a probability.
-\texttt{damegender} is based on open databases from official census and uses Machine Learning to guess strings not classified as names, such as diminutives or nicknames.
+Given the large amount of publicly available data on the software development process that can be retrieved and analyzed from the Internet (e.g., GitHub, StackOverflow), the importance of having methods and tools that help to identify the participation of women within large amounts of data is desirable.
+% Method name-to-gender inference
+In this paper we present a free software tool, called \texttt{damegender}, which we have conceived to infer the gender from the name of a participant.
+\texttt{damegender} is based on open databases from official census and uses Machine Learning.
+An experimental branch is currently under development for those cases where instead of names we have diminutives or nicknames.
 % Results 
-We have compared \texttt{damegender} with other tools, obtaining good results.
+We have compared \texttt{damegender} with other name-to-gender inference tools, obtaining good results.
 % Conclusions.
-We hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including gender gap) and diversity in the IT, and in particular in the free software community.
+We hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including the gender gap) and diversity in the IT, and in particular in the free software community.
 \end{abstract}
@@
 In recent times, many research investigations have been made on gender diversity in the IT domain.
+Examples of these efforts range from the participation in Twitter~\cite{burger2011discriminating,mislove2011understanding},
+in Wikipedia~\cite{antin2011gender,hill2013wikipedia}, in science~\cite{holman2018gender,dollar1999gender}, and more specifically in the software domain in StackOverflow~\cite{vasilescu2012gender,vasilescu2015gender}, GitHub~\cite{vasilescu2015gender}, and free software development~\cite{robles2014floss,izquierdo2018openstack,lee2019floss,catolino2019gender}.
 
-Examples of these efforts range from participation in Twitter~\cite{burger2011discriminating,mislove2011understanding},
-in Wikipedia ~\cite{antin2011gender,hill2013wikipedia}, in science~\cite{holman2018gender,dollar1999gender}, and more specifically in the software domain in StackOverflow~\cite{vasilescu2012gender}, GitHub~\cite{vasilescu2015gender} and in Free/Libre/Open Source Software development~\cite{robles2014floss}.
+The interest on gender diversity has become more and more relevant, and so does the identification methods that allow to perform comprehensive studies on gender representation, given the large amounts of data available, in particular from collaborative environments.
 
-The interest on gender diversity is become more and more relevant, and so does the identification methods that allow to perform comprehensive studies on gender representation in different domains, given the large amounts of data available, in particular from collaborative environments.
+The most used method to infer the gender of participants is by looking at their name.
+However, there are some studies about Twitter or GitHub were not only names to detect gender. 
+In them we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand-written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 
-There are different ways to detect gender from a person name and perhaps a surname.
-A first, more rudimentary, is based on data extracted from the census, Wikipedia, self-references in trust websites, searches in Google Images, among others.
-So, for instance, in some studies, for example, about Twitter or GitHub, some people have used not only names to detect gender. 
-Thus, we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand-written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
-
+Currently, there are different ways to detect gender from a person's name (and maybe other personal data, such as the surname, the geographic location, etc.).
+A first one is based on data extracted from the Wikipedia, self-references in trust websites, searches in Google Images, among others.
 Another way to do it is by using one of the existing Application Programming Interfaces (APIs).
-This paper is about the latter, about their possibilities and limitations.
-Therefore, (i) we evaluate the quality and price of different commercial solutions;
-(ii) we discuss about solutions using free licenses;
-(iii) we investigate what happens with those names without census, for example, nicknames or diminutives; and
-(iv) we elaborate on how massive gender detection from Internet, for example, mailing
-  lists or software repositories, can be done.
+This paper is about the latter, focusing on their possibilities and limitations.
+Therefore, (i) we evaluate the quality (and accessibility, including price) of different existing solutions;
+(ii) we discuss about their limitations; and
+(iii) we investigate what happens with those names not included in a census, for example, nicknames or diminutives.
+% and (iv) we elaborate on how massive gender detection from Internet, for example, mailing lists or software repositories, can be done.
 
@@
   detail; this tools has been compared with APIs using an international
-  dataset with good results.
+  dataset with good results; and
 
@@
 
-(iv) a proof-of-concept of \texttt{damegender} to detect
-  gender in mailing lists and software repositories.
+%(iv) a proof-of-concept of \texttt{damegender} to detect gender in mailing lists and software repositories.
 
@@
 
-In Section 2 we explain \texttt{damegender} as solution to the
-problem of name-to-gender inference. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
-Section 4 presents a feature comparison with other name-to-gender inference services and tools.
-Section 5 reports on values of accuracy and offers a confusion matrix using a scientific dataset.
-Section 6 is about how we use Machine Learning in \texttt{damegender}.
-Section 7 discusses limitations and further research, and concludes the paper.
+In Section~\ref{sec:damegender} we explain \texttt{damegender} as solution to the problem of name-to-gender inference.
+Section~\ref{sec:datasets} introduces the Open Datasets used as the \emph{golden truth}.
+Section~\ref{sec:comparison} presents a feature comparison with other name-to-gender inference services and tools.
+Section~\ref{sec:reproducing} reports on values of accuracy and offers a confusion matrix using a scientific dataset.
+Section~\ref{sec:ml} is about how we use Machine Learning in \texttt{damegender}.
+Section~\ref{sec:limitations} discusses limitations and further research, and concludes the paper.
 
 \section{Damegender}
+\label{sec:damegender}
 
-\texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool under a Free Software license (in particular, the GNU General Public License v3.0). 
-It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLTK) for Natural Language Processing~\cite{loper2002nltk}, Scikit for Machine Learning~\cite{pedregosa2011scikit}, Numpy for Numerical Computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
+\texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool released under a free software license (in particular, the GNU General Public License v3.0). 
+It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLTK) for natural language processing~\cite{loper2002nltk}, Scikit for machine learning~\cite{pedregosa2011scikit}, Numpy nor mumerical computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
 At its current point it is linked to Perceval~\cite{duenas2018perceval}, a tool specialized in retrieving and gathering data from software repositories, such as git and mailing lists.
-\texttt{damegender} is a Python package that can be installed using PIP (the package installer for Python) from the console.
+\texttt{damegender} is a Python package that can be easily installed using PIP (the package installer for Python) from the console.
 
@@
 The main reason for developing \texttt{damegender} is that there are not many free software tools that help in the identification of gender. 
-Before \texttt{damegender}, only \texttt{Gender guesser}\footnote{https://github.com/lead-ratings/gender-guesser} offered a free software solution in this field~\cite{krawetz2006gender}, and the project has not been active for more than three years now.
+Before \texttt{damegender}, only \texttt{Gender guesser}\footnote{https://github.com/lead-ratings/gender-guesser} offered a free software solution in this field~\cite{krawetz2006gender}, and the project (i) has not been active for more than three years now, (ii) lies technologically well behind other solutions.
 The best contribution of \texttt{Gender guesser} is the dataset containing 48,528 names with a good classification by countries\footnote{\url{https://raw.githubusercontent.com/lead-ratings/gender-guesser/master/gender_guesser/data/nam_dict.txt}}.
@@
 \section{Datasets}
+\label{sec:datasets}
 
@@
 
-There are several sources to create a databases, being the most common:  
+There are several sources to create these databases, being the most common:  
 (1) a census published with a free license (open census way), 
@@
 
-In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom.
-The datasets provided by the software package is incrementing the speed retrieving data.
+In \texttt{damegender}, we are including Open Data census about names and gender, usually offered by governments, such as the ones from Spain (by means of INE.es, the Spanish National Statistics Institute), Uruguay, USA and United Kingdom.
+%The datasets provided by the software package is incrementing the speed retrieving data.
 
@@
 
-Hence, we are using the census approach as base of truth to distinguish if a name is male or female in a geographical area. 
-Generally, a name has a strong weight to determine if it is a male or a female on this way.
+Generally, a name has a strong weight to determine if it is a male or a female using this procedure.
 For instance, David is registered 365,196 times as male, but 0 times as female in the data offered by the Spain National Institute of Statistics.
@@
 For instance, Andrea would be considered a female name in Germany, but a male name in Italy.
-However, many countries do not provide Open Data census about gender and names.
+Hence, we are using the census approach as base of truth to distinguish if a name is male or female in a geographical area. 
+However, many countries do not provide Open Data census about gender and names, although our hope is that they will do in the near future.
 
@@
 We have evaluated to include data from the second option (datasets released with a free license). 
-For instance, Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
-Another example is \texttt{Gender Guesser} a good dataset for international names with different categories to define the probability. 
-The problem with these data is that we have observed that they do not have the quality of National Statistics Institutes.
+For instance, the Natural Language Tool Kit offers 8,000 labeled English names classified as male or female. 
+Another example is \texttt{Gender Guesser}, which provides a good dataset for international names with different categories to define the probability. 
+Although these datasets can be incorporated into \texttt{damegender}, the problem with them is that, in general, we have observed that they do not have the quality of National Statistics Institutes.
 
@@
 
-The third approach is based on the trust on commercial solutions, in the same way we trust search engines when we make searches in Internet.
-This is because commercial APIs can be seen just a black box, so we do not know where the data comes from and how it has been treated.
-As at this point, commercial APIs offer better results as other solutions, \texttt{damegender} gives the possibility to include data from them.
-Thus, it is possible to download JSON files from the main commercial name-to-gender inference API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
+The third approach is to create the dataset from existing, usually commercial solutions.
+Here we have to trust their results, in the same way we trust search engines when we make searches in Internet.
+This is because commercial APIs can be seen as a black box -- we do not know where the data comes from and how it has been treated.
+But, at least, we can measure its quality, as we will do next.
+So, we will see that --at this point-- commercial APIs offer better results than other solutions.
+That is why \texttt{damegender} gives the possibility to include data from them.
+It is possible to download JSON files from the main commercial name-to-gender inference API solutions (e.g., \texttt{genderapi}, \texttt{genderize}, \texttt{namsor}, \texttt{nameapi}) and use it as the dataset.
 There are certain uses that are currently only available in such tools.
@@
 
-As a final goal, we envision to build a free dataset with names and gender, that builds on top of \texttt{Gender Guesser} and that can be made available as Wikidata. 
+As a final goal, we envision to build a free dataset with names and gender, that builds on top of \texttt{Gender Guesser} and that can be made available as Wikidata with a free license. 
 Perhaps, to complete this work, we need to combine an automated with a manual process as described in~\cite{10.7717/peerj-cs.156}.
@@
 \section{Feature comparison with other tools}
+\label{sec:comparison}
 
@@
 
-Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+Santamaría and Mihaljevic~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
 The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
@@
 \section{Reproducing values of accuracy and confusion matrix}
+\label{sec:reproducing}
 
@@
 \section{Machine Learning}
+\label{sec:ml}
 
@@
  NLTK (Bayes)                        &    0.862 &     0.902 &   0.902 &    1.0  \tabularnewline
- Multinomial Navie Bayes             &    0.782 &     0.791 &   0.791 &    1.0  \tabularnewline
+ Multinomial Naive Bayes             &    0.782 &     0.791 &   0.791 &    1.0  \tabularnewline
  Tree                                &    0.764 &     0.821 &   0.796 &    1.0  \tabularnewline
@@
 \section{Limitations and further research}
+\label{sec:limitations}
 

### 2020-05-16 paper2sattose.tex
Reviewed and new version

@@
 \texttt{damegender}\footnote{https://github.com/davidam/damegender} is a gender detection tool released under a free software license (in particular, the GNU General Public License v3.0). 
-It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLTK) for natural language processing~\cite{loper2002nltk}, Scikit for machine learning~\cite{pedregosa2011scikit}, Numpy nor mumerical computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
+It has been implemented in Python to take advantage of many other free software tools used in the scientific domain, such as the Natural Language Toolkit (NLTK) for natural language processing~\cite{loper2002nltk}, Scikit for machine learning~\cite{pedregosa2011scikit}, Numpy nor numerical computation~\cite{van2011numpy}, and Matplotlib to visualize results~\cite{hunter2007matplotlib}.
 At its current point it is linked to Perceval~\cite{duenas2018perceval}, a tool specialized in retrieving and gathering data from software repositories, such as git and mailing lists.
@@
 The subscription price is given in euro.}
+\vspace{0.3cm}
 \label{table:comparison}
@@
 
-Standard commercial Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
+Standard commercial name-to-gender inference Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
 To express geolocalization the user can also give surnames, a country ISO code, or specify a language.
@@
 
-Santamaría and Mihaljevic~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
+Santamaría and Mihaljevi\'c~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
 The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).
 
-We have used this comparison framework and have extended it with other tools, including \texttt{damegender} and updated it to today.
+We have used this comparison framework and have updated it to the current situation and extended it with \texttt{damegender}.
 Results can be found in Table~\ref{table:comparison}.
+As it can be observed, commercial services have the advantage of owning large datasets, in the order of millions or even of billions of entries.
+The most advanced ones offer support for non-Latin alphabets.
+On the other side, they limit the amount of free requests that can be performed and their costs range from 7 euro to 150 euro per month.
 
@@
 
+In this section, we will analyze the quality of service of existing name-to-gender inference services and tools.
 There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, recall).
-We can see in  the confusion matrix to understand where the different tools succeed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
+We can also inspect the confusion matrix to understand where the tools and services succeed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
 
-Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offer 7,000 names applying these methods. 
+Santamaría and Mihaljevi\'c~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name; they offer 7,000 names that can serve as the \emph{golden set} to evaluate them. 
 In their dataset, gender is classified as male, female or unknown. 
 We have used this dataset, not considering the unknown variable, for our experiments.
+
 The results, using common information retrieval metrics, can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
 Accuracy is the ratio of correctly predicted observation to the total observations.
-It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects.
-So, for a sample with 85 negative and 15 positive values, classifying all values as negative in this case gives a score of accuracy of 0.85.
-In those cases, it is better to report other measures, such as the balanced accuracy (bACC), which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
+It should be noted that \emph{Accuracy} can be a misleading metric for imbalanced data sets, such as the ones that we usually have in software development projects~\cite{ghosh2002free}.
+This is because for a sample with 85 negative and 15 positive values, classifying all values as negative would give an accuracy score of 0.85.
+In those cases, it is better to report other measures, such as the balanced accuracy, which normalizes true positive and true negative predictions by the number of positive and negative samples~\cite{mower2005prep}.
+As Santamaría and Mihaljevi\'c's data does not come from the IT domain, we can use accuracy in our research.
 Precision is the fraction of relevant instances among the retrieved instances.
@@
 Precision and recall are sometimes used together in the F1 Score (or f-measure) to provide a single measurement for a system.
-As can be observed, Genderapi and Genderize obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
+As can be observed, \texttt{Genderapi} and \texttt{Genderize} obtain the best results -- although all solutions are close and reach results better than 0.8 for accuracy, except for \texttt{Gender Guesser}.
+\texttt{damegender} offers relatively good results.
+Given, on the other hand, that its focus is mainly to be used in the IT domain, where the gender is highly imbalanced, we are very positive about its precision, as accuracy is not that relevant as noted above.
 
@@
 
-We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
-Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
-The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the values of accuracy with respect the previous comparison.
-In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
-In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
-For \texttt{Genderize}, we obtain the same results. 
+We have performed a comparison using a confusion matrix for the software/tools under study (see Table~\ref{table:ConfusionMatrixTables}).
+These results are very similar to the ones reported in~\cite{10.7717/peerj-cs.156}, which offers confidence that we have done our comparison correctly.
+The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) improve the values of accuracy with respect to the previous comparison.
+In particular, \texttt{Genderapi} improves the results for \emph{undefined}.
+For \texttt{Genderize}, we obtain exactly the same results. 
+For \texttt{Genderguesser}, however, we obtain different results, which is to some extent not expected, because the software has not been modified for several years.
 \texttt{Nameapi}'s results is changing from male to female with more errors. 
 In \texttt{Namsor}, the results are similar. 
-\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.
+\texttt{damegender} is not guessing undefined because we predict with machine learning (using an SVC algorithm) if the string is not in the database.
 
@@
 
-In Table~\ref{table:ApisAndErrors} we observe the different measures for errors in the APIs.
-Error coded defines if the true is different than the guessed one. 
-Error coded without na defines if the true is different than the guessed one, but without undefined results.
-Error gender bias allows to understand if the error is bigger for guessing males than females or viceversa.
-The weighted error defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
+In Table~\ref{table:ApisAndErrors} we can see the different measures for errors in the APIs.
+\emph{Error coded} defines if the true is different than the guessed one. 
+\emph{Error coded without na} defines if the true is different than the guessed one, but without undefined results.
+\emph{Error gender bias} allows to understand if the error is bigger for guessing males than females or viceversa.
+The \emph{weighted error} defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
 The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
@@
  Bernoulli Naive Bayes               &    0.699 &     0.965 &   0.816 &    1.0  \tabularnewline
-
 \hline
@@
 
-Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
-The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
-Our classifier is binary (only male and female).
+Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used in our guessing. 
+The best results are given by Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
+It should be noted that as a result of using machine learning techniques, our classifier is binary (male and female), so no \emph{unknown} is given as output.
 
@@
 
-We have developed some experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
-To test our approach, we have selected some features of the names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
+We have developed some additional experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
+To test our approach, we have selected some features of names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
 The selection of the features was verified with Principal Component Analysis.
@@
 
-For instance, the letter a varies 0.2 from males to females for USA and Uruguay, and 0.1 from males to females for United Kingdom, Australia and Spain.
-The last letter a varies 0.5 from males to females for Australia and Spain, around 0.4 for USA and United Kingdom and 0.2 in
+For instance, the feature \emph{containing the letter a in the name} varies 0.2 from males to females for USA and Uruguay, and 0.1 from males to females for United Kingdom, Australia and Spain.
+The feature \emph{the last letter is an a} varies 0.5 from males to females for Australia and Spain, around 0.4 for USA and United Kingdom and 0.2 in
 Uruguay.
-The last letter o from females to males varies 0.2 in (Spain, Australia) and is equal in Uruguay, USA, United Kingdom.
-For the last letter consonant all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
-So, last letter vocal is reverse than last letter consonant. 
-First letter consonant or first letter vocal is a non-significant feature due to offering similar results in English and Spanish.
+The feature \emph{the last letter is an o} varies 0.2 in (Spain, Australia) from females to males and is equal in Uruguay, USA, United Kingdom.
+For the feature \emph{the last letter is a consonant} all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
+So, the feature \emph{the last letter is a vocal} is, as expected, the reverse of the feature \emph{the last letter is a consonant}. 
+The features \emph{the first letter is a consonant} or \emph{the first letter is a vocal} are, however, non-significant as they offer similar results in English and Spanish.
 
-We have done this experiment with the NLTK and INE datasets, with the values of accuracy reaching up to 0.745.
+We have done this experiment with the NLTK and INE.es datasets, with the values of accuracy reaching up to 0.745.
 So it makes sense to expect better results in random datasets if we add new languages and countries.
@@
 However, their inner working cannot be studied (i.e., they work as a black-box for the outsider) and the fees that have to be paid for using their service are sometimes out of the reach of many researchers.
-That is why we propose a new tool, called \texttt{damegender}, with the aim of having open data on name-to-gender inference, which offers more flexibility and where researchers can build on top of it.
+That is why we propose a new tool, called \texttt{damegender}, with the aim of having open data of name-to-gender inference, which offers more flexibility and where researchers can build on top of it.
 This tool is offered under a free software license and is available on GitHub for download and enhancement.
@@
 
-In addition, we have shown a glimpse of how several features of the names could be used to guess the gender if we do not have the real name, but nicknames or diminutives.
-This experiment is at this point very preliminary, and we would like to work further on it.
+In addition, we have shown a glimpse of how several features of the names could be used to guess the gender if we do not have the real name, but nicknames or diminutives, a situation that is very common in free software development-supporting tools, such as IRC channels.
+This experiment is at this point very preliminary, and we would like to work more on it.
 
@@
 
-We would like to thank: Lucía Santamaría and Helena Mihaljevic for the previous work; Daniel Izquierdo and Laura Arjona for starting this research field at URJC; Jesus González Boticario and the people who work with him at UNED for the motivation towards the machine learning; all people working with Jesus González Barahona and Gregorio Robles.
+We would like to thank: Lucía Santamaría and Helena Mihaljevi\'c for the previous work; Daniel Izquierdo and Laura Arjona for starting this research field at URJC; Jesus González Boticario and his team at UNED for the motivation towards machine learning; all those working with Jesus González Barahona and Gregorio Robles.
 

### 2020-07-19 paper2sattose.tex
paper2sattose.tex: replace Namsor by Namsor (v1)

@@
 %Context
-Diversity in software development teams has been identified as one of the main ingredients of a more productive, more healthy software community.
+Diversity in software development teams has been identified as one of the main ingredients of a more productive and healthy software community.
 Thus, the interest of the research community in identifying who is contributing to a project has increased in the last years.
@@
 % Conclusions.
-We hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including the gender gap) and diversity in the IT, and in particular in the free software community.
+We hope the \texttt{damegender} tool can become a cornerstone for the scientific advancement of the study of gender (including the gender gap) and diversity in Information Technology (IT), and in particular in the free software community.
 \end{abstract}
@@
 
-The interest on gender diversity has become more and more relevant, and so does the identification methods that allow to perform comprehensive studies on gender representation, given the large amounts of data available, in particular from collaborative environments.
+The interest on gender diversity has become more and more relevant, and so does the identification methods that allow us to perform comprehensive studies on gender representation, given the large amounts of data available, in particular from collaborative environments.
 
 The most used method to infer the gender of participants is by looking at their name.
-However, there are some studies about Twitter or GitHub were not only names to detect gender. 
-In them we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand-written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
+However, there are some studies about Twitter and GitHub where not only names are used to detect gender. 
+In those studies we can find gender detection tools that infer the gender from faces in images~\cite{ranjan2017hyperface}, from hand-written annotations~\cite{liwicki2011automatic}, or from speeches~\cite{koppel2002automatically}.
 
@@
 Therefore, (i) we evaluate the quality (and accessibility, including price) of different existing solutions;
-(ii) we discuss about their limitations; and
+(ii) we discuss their limitations; and
 (iii) we investigate what happens with those names not included in a census, for example, nicknames or diminutives.
@@
   provides by the states built to understand current technologies in
-  detail; this tools has been compared with APIs using an international
+  detail; this tool has been compared with APIs using an international
   dataset with good results; and
@@
 (iii) a machine learning solution to strings not found in the census
-  dataset to approach the problem with nicknames and diminutives; and
+  dataset to approach the problem with nicknames and diminutives;
 
@@
 
-In \texttt{damegender}, we are including Open Data census about names and gender, usually offered by governments, such as the ones from Spain (by means of INE.es, the Spanish National Statistics Institute), Uruguay, USA and United Kingdom.
+In \texttt{damegender}, we are including Open Data census about names and gender, usually offered by governments, such as the ones from Spain (by means of INE.es, the Spanish National Statistics Institute), Uruguay, USA and UK.
 %The datasets provided by the software package is incrementing the speed retrieving data.
@@
 
-Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
-In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.
+Some Open Datasets, such the one offered by INE.es or the government of the US offer support for surnames and how they are related to ethnicity. 
+In particular, the dataset from the government of the US offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.
 
@@
 
-Standard commercial name-to-gender inference Application Programming Interfaces (APIs) usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
+Standard commercial name-to-gender inference APIs usually guess the gender for a single name or a list of names (from a CSV file or an API call). 
 To express geolocalization the user can also give surnames, a country ISO code, or specify a language.
@@
 The most advanced ones offer support for non-Latin alphabets.
-On the other side, they limit the amount of free requests that can be performed and their costs range from 7 euro to 150 euro per month.
+However, they limit the amount of free requests that can be performed and their costs range from 7 euro to 150 euro per month.
 
@@
 Damegender (SVC)\footnote{SVC is the acronym of Support Vector Classification, the Machine Learning algorithm that Damegender was using with this results} & 0.879 & 0.972 & 0.972 & 1.0\tabularnewline
-Namsor & 0.867 & 0.973 & 0.924 & 1.0\tabularnewline
+Namsor (v1) & 0.867 & 0.973 & 0.924 & 1.0\tabularnewline
 Nameapi & 0.830 & 0.974 & 0.905 & 1.0\tabularnewline
@@
 
-In this section, we will analyze the quality of service of existing name-to-gender inference services and tools.
-There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, recall).
-We can also inspect the confusion matrix to understand where the tools and services succeed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that appear.
+In this section, we will analyze the quality of the service of existing name-to-gender inference services and tools.
+There are different ways to express the probability of a successful identification (e.g., confidence, scale, accuracy, precision, and recall).
+We can also inspect the confusion matrix to understand where the tools and services succeed or fail, and to analyze the different errors measures (error coded, error coded without not applicable values, error gender bias, not applicable coded) that can happen.
 
 Santamaría and Mihaljevi\'c~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name; they offer 7,000 names that can serve as the \emph{golden set} to evaluate them. 
-In their dataset, gender is classified as male, female or unknown. 
+In their dataset, gender is classified as male, female, or unknown. 
 We have used this dataset, not considering the unknown variable, for our experiments.
@@
 \texttt{damegender} offers relatively good results.
-Given, on the other hand, that its focus is mainly to be used in the IT domain, where the gender is highly imbalanced, we are very positive about its precision, as accuracy is not that relevant as noted above.
+Given, however, that its focus is mainly to be used in the IT domain, where the gender is highly imbalanced, we are very positive about its precision, as accuracy is not that relevant as noted above.
 
@@
 Damegender        & male    & 3663 & 147  &   0 \tabularnewline
-(SVC)\footnotemark[1] & female  & 551  & 1497 &   0 \tabularnewline
+                  & female  & 551  & 1497 &   0 \tabularnewline
 Genderguesser     & male    & 3326 &  139 & 346 \tabularnewline
                   & female  & 78   & 1686 & 204 \tabularnewline
-Namsor            & male    & 3325 & 139  & 346 \tabularnewline
+Namsor (v1)       & male    & 3325 & 139  & 346 \tabularnewline
                   & female  & 78   & 1686 & 204 \tabularnewline
@@
 These results are very similar to the ones reported in~\cite{10.7717/peerj-cs.156}, which offers confidence that we have done our comparison correctly.
-The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) improve the values of accuracy with respect to the previous comparison.
+The most important tools (\texttt{Namsor (v1)}, \texttt{Genderapi}, and \texttt{Genderize}) improve the values of accuracy with respect to the previous comparison.
 In particular, \texttt{Genderapi} improves the results for \emph{undefined}.
@@
 \texttt{Nameapi}'s results is changing from male to female with more errors. 
-In \texttt{Namsor}, the results are similar. 
+In \texttt{Namsor (v1)}, the results are similar. 
 \texttt{damegender} is not guessing undefined because we predict with machine learning (using an SVC algorithm) if the string is not in the database.
@@
 Genderize & 0.276 & 0.261 & 0.0204 & -0.0084 \tabularnewline 
-Namsor & 0.332 & 0.262 & 0.095 & 0.01 \tabularnewline
+Namsor (v1) & 0.332 & 0.262 & 0.095 & 0.01 \tabularnewline
 Nameapi & 0.361 & 0.267 & 0.129 & 0.001 \tabularnewline
@@
 
-In Table~\ref{table:ApisAndErrors} we can see the different measures for errors in the APIs.
-\emph{Error coded} defines if the true is different than the guessed one. 
-\emph{Error coded without na} defines if the true is different than the guessed one, but without undefined results.
-\emph{Error gender bias} allows to understand if the error is bigger for guessing males than females or viceversa.
-The \emph{weighted error} defines if the true value is different than the guessed one, but giving a weight to the guessed as undefined.
-The most relevant information is a high index of errors for \texttt{Nameapi} and \texttt{Namsor},
-while \texttt{GenderApi} and \texttt{damegender} have a low index of errors.
+In Table~\ref{table:ApisAndErrors} we can see the different measures
+for errors in the APIs.  \emph{Error coded} defines if the true is
+different than the guessed one.  \emph{Error coded without na} defines
+if the true is different than the guessed one, but without undefined
+results.  \emph{Error gender bias} allows to understand if the error
+is bigger for guessing males than females or viceversa.  The
+\emph{weighted error} defines if the true value is different than the
+guessed one, but giving a weight to the guessed as undefined.  The
+most relevant information is a high index of errors for
+\texttt{Nameapi} and \texttt{Namsor (v1)}, while \texttt{GenderApi}
+and \texttt{damegender} have a low index of errors.
 
@@
   \hline
-Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
+Dataset & contains a & last is a & last is o & last is consonant & last is vowel & 1st is consonant & 1st is vowel  \tabularnewline
   \hline
@@
 We have developed some additional experimental functionality that allows to analyze our database according to some features using machine learning algorithms.
-To test our approach, we have selected some features of names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vocal, last letter is a vocal, last letter is a consonant, or last letter is a.
+To test our approach, we have selected some features of names, such as a being the first letter, a (or o) being the last letter, contains the letter a, first letter is a vowel, last letter is a vowel, last letter is a consonant, or last letter is a.
 The selection of the features was verified with Principal Component Analysis.
@@
 
-For instance, the feature \emph{containing the letter a in the name} varies 0.2 from males to females for USA and Uruguay, and 0.1 from males to females for United Kingdom, Australia and Spain.
+For instance, the feature \emph{containing the letter a in the name} varies 0.2 from males to females for USA and Uruguay, and 0.1 from males to females for United Kingdom, Australia, and Spain.
 The feature \emph{the last letter is an a} varies 0.5 from males to females for Australia and Spain, around 0.4 for USA and United Kingdom and 0.2 in
@@
 For the feature \emph{the last letter is a consonant} all countries give as a result that males do have it more frequently, with results that range from 0.3 to 0.5: Uruguay and USA (0.5), United Kingdom (0.4), Australia and Spain (0.3).
-So, the feature \emph{the last letter is a vocal} is, as expected, the reverse of the feature \emph{the last letter is a consonant}. 
-The features \emph{the first letter is a consonant} or \emph{the first letter is a vocal} are, however, non-significant as they offer similar results in English and Spanish.
+So, the feature \emph{the last letter is a vowel} is, as expected, the reverse of the feature \emph{the last letter is a consonant}. 
+The features \emph{the first letter is a consonant} or \emph{the first letter is a vowel} are, however, non-significant as they offer similar results in English and Spanish.
 
